---
title       : Performance drivers
description : In addition to studying portfolio performance based on the observed portfolio return series, it is relevant to find out how individual (expected) returns, volatilities and correlations interact to determine the total portfolio performance. 

--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:27fc122215
## The sources of portfolio performance

The portfolio mean return and volatility depends on the mean return and volatility of the assets in which the portfolio
is invested. In the videos, I will show you the exact nature of the dependence. But let us first get a feeling for this dependence by analyzing the monthly returns of an investment in US equities and an investment in US bonds.  

As a reference, we take a 60/40 portfolio that invests every month 60% in equities and 40% in bonds. The individual returns are available in the workspace as `returns_bonds` and `returns_equities`, while the portfolio returns are avalailble as  `returns_6040`.   
(the ETF. 

Verify in the R-console which of the following statements about the relationship between portfolio performance and the individual assets' performance is false. 


*** =instructions
- A portfolio is less volatile than the sum of its component volatilities. This implies that the standard deviation of the 60/40 portfolio is less than `0.6*sd(returns_equities)+0.4*sd(returns_bonds)`. We call this the subadditivity property of the standard deviation.    
- The portfolio mean return equals the position weighted sum of the component mean returns, that is, `mean(returns_6040)==0.6*mean(returns_equities)+0.4*mean(returns_bonds)`.  
- None of the above statements is correct. 

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])

### plot
weights <- c(0.6,0.4)
prices <- cbind(eq_prices,bond_prices)
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_6040 <- Return.portfolio(returns, weights,rebalance_on="months")
# check: all.equal( as.numeric(returns_6040),as.numeric(0.6*returns[,1]+0.4*returns[,2])  )
prices_6040 <- zoo(cumprod(1+returns_6040),order.by=time(returns_6040))
temp <- c( as.numeric(eq_prices), as.numeric(bond_prices) )
ylim <- c( min( temp ) , max(temp) )
plot.zoo(eq_prices , main="Equity-bond portfolios",ylab="", xlab="",ylim=ylim)
lines(bond_prices,col="red")
lines(prices_6040,col="blue")
legend("topleft", legend=c("Equities (ticker: SPY)","Bonds (ticker: AGG)","60/40 Equities-Bonds") , col=c("black","red","blue") , lwd = 1 )

###
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```

--- type:NormalExercise lang:r xp:50 skills:1     key:69c5664fdb
## Sensitivity of the Sharpe ratio to the porfolio weights

The portfolio mean return and standard deviation depends on the mean return and standard deviation of the underlying
asset returns, as well as the portfolio weights. 

In the special case of investing in two assets, namely US equities and US bonds, we can easily find the weight that would produce the highest Sharpe ratio. One way to do so is the brute force approach of trying a large number of possible weights and keeping the weight that yields the highest value for hte portfolio Sharpe ratio. Since the portfolio is fully invested, the weigths sum to one and we thus need to determine only one of the two weights. Let's take the weights of the equity investment as the variable to be optimized. 



*** =instructions
- Complete the R script that runs a brute force search to determine the portfolio weights for which the Sharpe ratio is the highest, assuming a zero risk free rate.  

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sample_code
```{r}
# returns_equities and returns_bonds are preloaded

# grid defining possible weights equities 
grid <- seq(0,1,0.01)
print(grid);
# initialize vector in which the Sharpe ratios will be saved
vsharpe <- rep(NA,length(grid))
# initialize counter
i <- 1
#loop over the elements in the search grid
for(w in grid){
  # portfolio returns are monthly rebalanced
  preturns <- w*returns_equities+(1-w)*returns_bonds
  # compute the shapre ratio of preturns
  vsharpe[i] <- 
  # update counter  
  i <- 
}
# plot the obtained Sharpe ratios against the equity weights
plot(  ,   ,xlab="Equity weights",ylab="Ann. Sharpe ratio")
abline(v=grid[vsharpe==max(vsharpe)],lty=3)
```


*** =solution
```{r}
# PerformaneAnalytics is pre-loaded
# returns_equities and returns_bonds are preloaded
grid <- seq(0,1,0.01)
vsharpe <- rep(NA,length(grid))
i <- 1
for(w in grid){
  preturns <- w*returns_equities+(1-w)*returns_bonds
  vsharpe[i] <- SharpeRatio.annualized(preturns)
  #vsharpe[i] <- SharpeRatio.annualized(preturns)
  i <- i+1
}
plot(grid,vsharpe,xlab="weights",ylab="Ann. Sharpe ratio")
abline(v=grid[vsharpe==max(vsharpe)],lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:bf1fc7dc803e342b4475c296bbebc2bab9b26f32
## The drivers in case of two assets

*** =video_link
//player.vimeo.com/video/108225030


--- type:NormalExercise lang:r xp:100 skills:1 key:b6a986b0e53bd119933fc39df94f21f2bf1e4318
## Interpreting correlation

The correlation thus tells us the tendence of two asset returns to move together. In the extreme case of two identical asset returns, the correlation will be one and there is no diversification potential. In the other extreme case where, if one asset return is above average, the other one is almost always below average, the correlation is negative. It is zero, when the asset returns are independent. 



Measuring is knowing. Let us now compute the correlation between equity returns and bond returns. We will do this first taking all returns in the sample. However, just like volatilities, these correlations are dynamic, as we will see from the time series of correlations that we obtain when estimating the correlation over rolling samples. 

*** =instructions
- Plot the equity returns (`returns_equities`) against the bond returns (`returns_bonds`) using the function [chart.Scatter](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.Scatter) with the bond returns on the x-asis. Do you see a relation? 
- Compute first the correlation between the variables `returns_equities` and `returns_bonds` using the standard function [cor](). 
- Repeat this using the function [chart.Correlation](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.Correlation), with as argument the multivariate time series `returns`. 
- Commpute the rolling 24-month estimates of the bond-equity correlation using the function [chart.RollingCorrelation](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.RollingCorrelation). 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# load package
library(PerformanceAnalytics)
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sample_code
```{r}
# scatter plot

# calculation using cor

# merge returns_equities and returns_bonds into returns

# calculation and visualization using chart.Correlation

# rolling estimates using chart.RollingCorrelation

```

*** =solution
```{r}
# scatter plot
chart.Scatter(returns_equities,  returns_bonds)
# calculation using cor
cor(returns_equities,  returns_bonds)
# merge returns_equities and returns_bonds into returns
returns <- merge(returns_equities,returns_bonds)
# calculation and visualization using chart.Correlation
chart.Correlation(returns)
# rolling estimates using chart.RollingCorrelation
chart.RollingCorrelation(returns_equities, bonds,returns_equities,width=24)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1     key:f4f41f6b81
## The impact of correlation on performance.

The correlation is an important driver of portfolio performance in terms of avoiding high risks. In fact,the lower the correlation is, the more the risks in the indvidual investments offset each other. Those gains in diversification then lead to a lower portfolio volatility. 

This diversifciation impact of decreasing correlation on portfolio volatility is illustrated in your wokspace.
The plot shows the annualized volatility of the 60/40 and the 40/60 portfolio invested in equities and bonds for various values of the correlation between -1 and 1. 

Let us now test whether your understanding of changes in correlation on portfolio volatility is correct. Suppose there is  an equally weighted portfolio of two assets. Their correlation jumps from 0 to 0.5. Then the portfolio variance:    

*** =instructions
- Decreases
- Increases
- Does not change  

*** =hint
Remember, from the video, that the portfolio variance equals  the squared weight of asset 1 times the variance of asset 1, plus the squared weight of asset 2 times the variance of asset 2, plus two times the product of the weights, the volatilities and their correlation. 

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# load package
library(PerformanceAnalytics)
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities

# standard deviations
sig1 <- sqrt(12)*sd(returns_equities)
sig2 <- sqrt(12)*sd(returns_bonds)
# corrleation grid
grid <- seq(-1,1,0.01)
# initializing solution vector
w <- 0.6
vpsd_6040 <- sqrt( w^2*sig1^2 + (1-w)^2*sig2^2 + 2*w*(1-w)*sig1*sig2*grid  )
w <- 0.4
vpsd_4060 = sqrt( w^2*sig1^2 + (1-w)^2*sig2^2 + 2*w*(1-w)*sig1*sig2*grid  )  
plot(grid,vpsd_6040,xlab="correlation",ylab="Ann. standard deviation",
      ylim = c(sig2,sig1),type="l")
lines(grid,vpsd_4060,col="red")
legend("topleft", legend=c("60/40 Equities-bonds","40/60 Equities-Bonds") , col=c("black","red") , lwd = 1 )


 # ((0.5^2*0.10^2+0.05^2*0.20^2+2*0.5*0.5*0.5)-(0.5^2*0.10^2+0.05^2*0.20^2))/(0.5^2*0.10^2+0.05^2*0.20^2)
```


*** =sct
```{r}
# sct code
test_mc(2) 
```

--- type:VideoExercise lang:r xp:50 skills:1 key:a2866e000af4e20577dc0d48d46765e2da1a7d9b
## The general case using matrix notation

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:080c6863fc
## Making a risk-reward scatter diagram

A realistic portfolio is not only invested in US equities and US bonds. It also seeks for diversification by investing in real estate and commodities, among others. We therefore extend our investment opportunity consisting of the US equities ETF (SPY) and US bonds ETF (AGG) to include also a real estate investment trust (VEIEX) and an ETF tracking the GSCI commodities index (GSG).

The monthly returns on those investment are available as the variable `returns` in your workspace. 

A good way to visualize the relative attractivenss of investments in terms of risk and rewards is to make a scatter 
plot of the average returns against the portfolio volatilities. 

We need to compute those average returns and volatilities for each asset, corresponding to a column in `returns`. These calculations are made easy using the function `apply` with as first argument the return data file, the second argument is 2 indicating the calculation should be column-wise and the third argument is the name of the function that needs to be applied on each column. 

*** =instructions
- Compute the vector of average returns on those four investments using the R command `apply(returns,2,"mean")` and call this `vmeans`.
- Do the same to compute the vector standard deviations and call this `vsd`.
- Use the following R instructions to produce the scatter plot of average returns with respect to standard deviations:
`plot(vsd,vmeans,col="gray",xlim=c(0,0.10))
text( vsd ,  vmu ,  labels=colnames(returns), cex= 0.7)`
- Add a dashed horizontal line at `h=0` using [abline](http://www.rdocumentation.org/packages/graphics/functions/abline) by setting the argument `lty=3`
*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
# vector of means
 
# vector of standard deviations

# scatter plot
 


```


*** =solution
```{r}
# vector of means
vmeans <- apply(returns,2,"mean")
# vector of standard deviations
vsd <- apply(returns,2,"sd")
# scatter plot
plot(vsd,vmeans,col="gray",xlim=c(0,0.10))
text( vsd , vmeans,  labels=colnames(returns), cex= 0.7)
abline(h=0,lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 
## The covariance matrix

The assets' volatitilities reflect the individual risks of the asset. What matters for the portfolio investors is however the portfolio risk caused by each of the assets, which depends also on the return volatilities.  

Let us compute and analyze the covariance and correlation matrix on the monthly returns of the four asset classes. 

The monthly returns on the investments are available as the variable `returns` in your workspace, together with the previously computed vector of standard deviations `vsd`. 


*** =instructions
- Compute the covariance matrix;
- Compute the correlation matrix;
- Verify that the covariance between bond returns and equity returns equals the product between their standard deviations and correlation. 
- Do the same to compute the vector standard deviations and call this `vsd`.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
# vector of means
 
# vector of standard deviations

# scatter plot
 


```


*** =solution
```{r}
# vector of means
vmeans <- apply(returns,2,"mean")
# vector of standard deviations
vsd <- apply(returns,2,"sd")
# scatter plot
plot(vsd,vmeans,col="gray",xlim=c(0,0.10))
text( vsd , vmeans,  labels=colnames(returns), cex= 0.7)
abline(h=0,lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1   key:21e408b8c65335ed2599fcb8ff64b5c9b352c374
## Portfolio performance and matrix formula's in practice (to do)

A realistic portfolio is not only invested in US equities and US bonds. It also seeks for diversification by investing in real estate and commodities, among others. We therefore extend our investment opportunity consisting of the US equities ETF (SPY) and US bonds ETF (AGG) to include also a real estate investment trust (VEIEX) and an ETF tracking the GSCI commodities index (GSG).

The monthly returns on those investment are available as the variable `returns` in your workspace. 

*** =instructions
- Use the function [Return.portfolio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.portfolio) to compute the time series of returns for the monthly rebalanced portfolio that is 40% invested in equities, 40% in bonds, 10% in real estate and 10% in commodities. Call this new variable `preturns`.
- Merge `returns` and `preturns` and apply the function [table.AnnualizedReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/table.AnnualizedReturns) to the merged object in order to compare the portfolio performance. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
re_prices <- re_prices/as.numeric(re_prices[1])
comm_prices <-  comm_prices/as.numeric(comm_prices[1])


### plot

prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
returns <- Return.calculate(prices)
returns <- returns[(-1),]
temp <- c( as.numeric(eq_prices), as.numeric(bond_prices),as.numeric(re_prices), as.numeric(comm_prices)   )
ylim <- c( min( temp ) , 1.4*max(temp) )
plot.zoo(eq_prices , main="Performance across asset classes",ylab="", xlab="",ylim=ylim)
lines(bond_prices,col="red")
lines(re_prices,col="blue")
lines(comm_prices,col="purple")
legend("topleft", legend=c("Equities (ticker: SPY)","Bonds (ticker: AGG)", "Real estate (ticker: VEIEX)",
                           "Commodities (ticker: GSG)") , col=c("black","red","blue","purple") , lwd = 1 ,ncol=1)

###
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
weights <-  
preturns <- Return.portfolio(returns, weights,rebalance_on="months")
# merge
 
# compare annualized performance
table.AnnualizedReturns(returns_merged) 

```

*** =solution
```{r}
# sample code
weights <- c(0.4,0.4,0.1,0.1)
preturns <- Return.portfolio(returns, weights,rebalance_on="months")
# merge
returns_merged <- merge(returns,preturns)
# compare annualized performance
table.AnnualizedReturns(returns_merged)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```






--- type:NormalExercise lang:r xp:100 skills:1 key:9786a14448b7689cc1be071bd4f4408ceadd0ee7
## Portfolio mean return and volatility using matrix notation

Use cov to compute the sample covariance matrix. use of %*% or StdDev in PerformanceAnalytics. to see.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1   key:ba9b12d382
## The portfolio's risk budget

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:0931e83c3a402b2fb2d78958981928a327428e4d
## Whodunit

compute the portfolio risk budget... barplot? function [StdDev](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/StdDev)

*** =instructions
- Set `vol_budget <- StdDev(returns,portfolio_method="component",weights=c(0.4,0.4,0.1,0.1))` and print this variable;
- Combine the weights and the percentage risk contributions in a table called `weights_percrisk`;
- Make a barplot and note how different the percentage risk contributions are compared to the portfolio weights. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code

vol_budget <- StdDev(returns,portfolio_method="component",weights=c(0.4,0.4,0.1,0.1))
out <- cbind( c(0.4,0.4,0.1,0.1) ,  vol_budget$pct_contrib_StdDev)
colnames(out) <- c( "weights" , "perc risk contrib")
barplot(out)
 
  
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

