---
title       : Optimizing the portfolio 
description : We have up to now considered the portfolio weights as given. In this chapter you learn how to determine in R the portfolio weights that are optimal in terms of achieving a target return with minimum risk, while satisfying constraints on the portfolio weights.  
 
--- type:VideoExercise lang:r xp:50 skills:1 key:2d45c303bc2e49767a9cd744f15a31938a2ec7d9
## Modern portfolio theory of Harry Markowitz


*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v1/hls-ch4_1.master.m3u8

--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:139d7eabbd827860709b29f2a8d968614d1e3cb0
## Mean-variance based investing in DJIA stocks 

In this chapter you will explore portfolio optimization. For the upcoming exercises you will focus on the Dow Jones Industrial Average (DJIA) stocks. Like the S&P 500 portfolio studied in Chapter 2, the Dow Jones Industrial Average portfolio is an important reference portfolio for the US equity market. It is invested in 30 large publicly owned companies based in the United States.

You will be using the sample of monthly returns from January 1991 until December 2015. These return data are available in the workspace as the variable `returns`.   

Assume you an investor that likes high returns with minimum risk. Which of the following statements is false:

*** =instructions
- You would rather invest all of your capital in Home Depot (ticket :HD) than in Caterpillar (ticker: CAT) because Home Depot has a higher average return, and lower risk.
- By combining investments, it may be possible to obtain a portfolio with a lower variance than the variance of the Exxon Mobil Corporation's stock (ticker: XOM). 
- If an investor is not allowed to borrow, then there is no portfolio invested in thirty Dow Jones Industrial Average stocks, offering a higher average return than the one of Apple (ticker: AAPL).  
- There is a total of 300 monthly observations in our sample of monthly returns. The highest return is around 88% while the lowest return is around -58%. 

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices <- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

stockmu = apply( X = returns , MARGIN = 2 , FUN="mean")
stocksd = apply( X = returns , MARGIN = 2 , FUN="sd")

# make the scatter plot and add tickers

plot(stocksd,stockmu,col="gray" , xlab="Standard deviation (monthly)", ylab="Average return (monthly)",xlim=c(0,1.1*max(stocksd)),ylim=c(0,1*1*max(stockmu)))
text( stocksd[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))] , 
       stockmu[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))] ,  
      labels=colnames(returns )[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))], cex= 0.7)
text( stocksd[c("HD","CAT","XOM","AAPL")]  ,  stockmu[c("HD","CAT","XOM","AAPL")] ,  labels=  c("HD","CAT","XOM","AAPL") , cex= 0.9,col="red")
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:NormalExercise lang:r xp:100 skills:1   key:8c67f73515
## Exploring monthly returns of the 30 DJIA stocks 

The 1991-2015 monthly returns on the 30 DJIA stocks are available in the workspace as the variable `returns`. This exercise will help you get comfortable with the data that you will be using for the remainder of the exercises.


*** =instructions
- Verify that `returns` is an object of the xts-class using the function `class()`.
- Create a vector of row means of `returns` using the function `rowMeans()`. Assign this to `row_means`.
- Plot `row_means` using `plot.zoo()`.
- Create a correlation matrix of the returns called `z`.
- Plot the histogram of the correlations between the monthly returns of the DJIA stocks. Because of the symmetry of the correlation matrix, you only need the lower triangular elements, which, for any square matrix `z`, you can select using `z[lower.tri(z)]`.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# Verify the class of returns 


# Create a vector of row means


# Plot row_means


# Create a correlation matrix


# Plot the histogram of the correlation matrix

```

*** =solution
```{r}
# Verify the class of returns 
class(returns)

# Create a vector of row means
row_means <- rowMeans(retrns)

# Plot row_means
plot.zoo(row_means)

# Create a correlation matrix
z <- cor(returns)

# Plot the histogram of the correlation matrix
hist(z[lower.tri(z)])
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:VideoExercise lang:r xp:50 skills:1 key:b48ce393f1
## Modern portfolio theory of Harry Markowitz


*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v2/hls-ch4_1.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:c066b2d5f610d13afdb48831249d18142590def7
## Finding the mean-variance efficient portfolio  

A mean-variance efficient portfolio can be obtained as the solution of minimizing the portfolio variance under the constraint that the portfolio expected return equals a target return. A convenient R function for doing so is the function [`portfolio.optim`](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim) in the R package [tseries](http://www.rdocumentation.org/packages/tseries). 
Its default implementation finds the mean-variance efficient portfolio weights under the constraint that the portfolio return equals the return on the equally-weighted portfolio. 
The only argument needed is the monthly return data on the portfolio components for which the weights need to be determined.


*** =instructions
- load the library `tseries`. 
- Create a mean-variance efficient portfolio of monthly returns using `portfolio.optim`, called `optim`. 
- Create a vector of optimum asset weights that are larger than or equal to 1%, call this `optim_weights`. Portfolio weights can be found in `optim$pw`. 
- Create a barplot of `optim_weights`.
- Print the expect portfolio return (`out$pm`) and volatility (`out$ps`) of the optimized portfolio. 
- Verify that the optimized portfolio returns in `out$px` dominate the returns of the equally weighted portfolio.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices <- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# Load tseries

# Create an optimized portfolio of returns


# Create a vector of optimum weights
optim_weights
names(optim_weights) <- colnames(returns)


optim_returns <-  out$px[out$px >= .01]
# round sorted
round( sort(    ,decreasing=T) ,4)
# barplot
   (optim_weights[optim_weights>=0.01])
# print expected return and volatility 

# equally weighted returns
ew_returns <- rowMeans(returns)
# strict lower standard deviation
 
# same or higher return
 
```

*** =solution
```{r}
#####josi progress####
# Load tseries
library(tseries)

# Create an optimized portfolio of returns
optim <- portfolio.optim(returns)

# Create a vector of optimum weights
optim_weights <- optim$pw[optim$pw >= 0.01]
names(optim_weights) <- colnames(returns)

##### end josi #######

# load tseries. 
library(tseries)
# optimize
out <- portfolio.optim(returns)
# output
optim_weights <- out$pw
names(optim_weights) <- colnames(returns)
optim_returns <- xts(out$px,order.by=time(returns))
# round sorted
round(sort(optim_weights,decreasing=T) ,4)
# barplot
barplot(optim_weights[optim_weights>=0.01])
# dominate
ew_returns <- rowMeans(returns)
# strict lower standard deviation
out$ps ; sd(ew_returns)
( out$ps < sd(ew_returns) )
# same or higher return
out$pm ; mean(ew_returns)
out$pm >= mean(ew_returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1  key:2240e86100
## Effect of the return target 

A more general specification of the function [portfolio.optim](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim) is as follows:
```
portfolio.optim(x, pm = mean(x),
                shorts = FALSE,  reshigh = NULL)
```

The argument ``pm`` sets the target return. The argument `reshigh` specifies the upper constraints on the portfolio weights.The argument ``shorts`` is a logical statement specifying whether negative weights are forbidden or not, by default `shorts=FALSE`. 

In this exercise you will quantify the risk-reward trade-off when choosing the mean-variance efficient portfolio. Note that the higher the target return, the higher the portfolio volatility. 

*** =instructions
- There is a isk-reward trade-off when choosing the mean-variance efficient portfolio. The higher the target return, the higher will be the corresponding portfolio volatility. When the target return is 10% higher than the average return, how much % increase is there in the portfolio volatility?  
- Compute the Sharpe ratio of the two portfolios when the monthly risk free rate is 0.1%. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

library(tseries)
```

*** =sample_code
```{r}
meanreturns = mean(returns)
# compute standard deviation when target retun is mean return
out1 <- portfolio.optim(returns, pm = meanreturns )
sd1 <-
 
# compute standard deviation when target retun is 1.1 times the mean return
out2 <-
sd2 <- 
# compute relative increase in standard deviation

```

*** =solution
```{r}
# solution code
# load tseries. 
library(tseries)
meanreturns <- mean(returns)
# compute standard deviation when target retun is mean return
out1 <- portfolio.optim(returns, pm = meanreturns )
sd1 <- out1$psd
 
# compute standard deviation when target retun is 1.1 times the mean return
out2 <- portfolio.optim(returns, pm = 1.1*meanreturns )
sd2 <- out2$psd
# compute relative increase in standard deviation
(sd2-sd1)/sd1
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6f3df5bc7125d675f226ddc4c69639e2cf24d23c
## Imposing weight constraints

Investors are often constrained by the maximum values allowed for the portoflio weights. This constraints can actually be an advantage. The advantage of a maximum weight constraint is that the subsequent portfolio will be less concentrated in certain assets. There is a disadvantage to this though. The disadvantage is that the same target return, may no longer be possible or will be obtained at the expense of a higher volatility. 

Finding the optimized portfolio under weight constraints is made easy with the `PerformanceAnalytics` package. The function [`portfolio.optim`](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim) allows you to set weight constraints within the `reshigh` argument. `reshigh` requires a vector of maximum weights for each asset. 

In this exercise you will create three portfolios with different maximum weight constraints. For this exercise it is important to know the output of the `portfolio.optim` function. This function creates a list containing four components:
  - `$pw`: the portfolio weights
  - `$px`: the returns of the overall portfolio
  - `$pm`: the exepcted return portfolio
  - `$ps`: the standard deviation of the portfolio returns

*** =instructions
- Create three vectors of maximum weights for each asset (column) in `returns` using the `rep()` function. The first vector will contain maximum weights of 100%, the second 10%, and the third 5%. Call these `weights1`, `weights2`, `weights3`, respectively. 
- Create an optimum portfolio with maximum weights of 100% called `optim1`.
- Create an optimum portfolio with maximum weights of 10% called `optim2`.
- Create an optimum portfolio with maximum weights of 5% called `optim3`.
- Calculate how many positions (assets) have a weight that is greater than 1% for each portfolio. Access weights by using `$pw` after the portfolio name.
- Print the volatilities of the portfolio you created. 

*** =hint
The argument `reshigh` needs to be a vector. 

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

library(tseries)
```

*** =sample_code
```{r}
# Create vectors of maximum weights


# Create an optimum portfolio with max weights of 100%


# Create an optimum portfolio with max weights of 10%


# Create an optimum portfolio with max weights of 5%


# Calculate how many assets have a weight that is greater than 1% for each portfolio


# Print portfolio volatilites 
```

*** =solution
```{r}
# Create vectors of maximum weights
weights1 <- rep(1, ncol(returns))
weights2 <- rep(.1, ncol(returns))
weights3 <- rep(.05, ncol(returns))

# Create an optimum portfolio with max weights of 100%
optim1 <- portfolio.optim(returns, reshigh = weights1)

# Create an optimum portfolio with max weights of 10%
optim2 <- portfolio.optim(returns, reshigh = weights2)

# Create an optimum portfolio with max weights of 5%
optim3 <- portfolio.optim(returns, reshigh = weights3)

# Calculate how many assets have a weight that is greater than 1% for each portfolio
sum(optim1 > .01)
sum(optim2 > .01)
sum(optim3 > .01)

# Print portfolio volatilites 
out1$ps
out2$ps
out3$ps
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:461436d2ee4b426f97b3c95b6ee39836ae38b34f
## The efficient frontier

*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v2/hls-ch4_2.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:2c03a87fc29fb66e35d8cdbf244b3dcf41305824
## Computing the efficient frontier using a grid of target returns

As we have seen, one approach to compute the efficient frontier is to first define the grid of target returns and then find, for each target return, the portfolio that has an expected return equal to the target return at the lowest possible variance. 

But what is a reasonble grid of target returns? We will set the maximum target return to the maximum average return of the stocks. Ideally, we set the minimum target return to the return of the minimum variance portfolio. Since we don't know this minimum variance portfolio return yet, let's therefore take the largest possible grid starting with the minimum average returns of all stocks. 


*** =instructions
- Complete the code specifying the grid of 50 equally spaced return targets starting at the minimum and maximum average return 
- Run the loop computing the optimized weights and corresponding mean return and volatility
- Plot the portfolio mean against the portfolio volatility of the optimized portfolios
- Improve the efficient frontier plot by removing the inefficient portfolios offering a lower return at a higher risk compared to the minimum variance portfolio.  

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
stockmu <- colMeans(returns)
grid <- seq( , ,length.out=50 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights <- c()
for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <- c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
colnames(mweights) <- colnames(returns)
# report the scatter plot of vpm agains vpsd
plot( ,  , xlab="(monthly) Volatility" , ylab="(monthly) average returns", 
      main="Minimum variance portfolios under a return target constraint",
      type="l",lwd=2)
```

*** =solution
```{r}
# solution code
stockmu <- colMeans(returns)
grid <- seq(min(stockmu),max(stockmu),length.out=50)  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights <- c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <- c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
# report the scatter plot of vpm agains vpsd
plot(vpsd,vpm, xlab="(monthly) Volatility" , ylab="(monthly) average returns", 
     main="Minimum variance portfolios \n under a return target constraint",type="l",lwd=2)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1    key:e647c251a7
## Properties of the efficient frontier

The curve shown in the previous plot shows the solution of minimizing the variance under the constraint that the portfolio expected return equals the target return. When the target return is below the minimum variance return, the obtained portfolio is not efficient because a higher return at a lower risk can be optained by investing in the minimum variance portfolio.
The optimized portfolios often are also not realistic because they tend to invest large weights in only a few assets. A practical solution to avoid this is to impose weight constraints. 

The effect of such a weight constraint is shown in the work space, where you can see the efficient frontier plots obtained for the DJIA stocks under a maximum 100% weight constraint (black line), a maximum 10% weight constraint (red line) and a maximum 5% weight constraint (blue). Which of the following statements is false.  

*** =instructions
- Imposing a more strict weight constraint, leads to a shift to the right of the efficient frontier.    
- The portfolios on the dashed lines are not mean-variance efficient. 
- The observed average return of any risky portfolio is always higher than the risk-free rate. 

*** =hint
hint

*** =pre_exercise_code
```{r}
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

stockmu <- apply( X = returns , MARGIN = 2 , FUN="mean")
stocksd <- apply( X = returns , MARGIN = 2 , FUN="sd")

grid <- seq(min(stockmu),max(stockmu),length.out=50 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns,pm=mutarget,
                            reshigh = rep(1,ncol(returns)) ),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <- c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}

# assign the stock names to the column names of mweights
colnames(mweights) <- colnames(returns)


##################################################################################
# To avoid the concentration, we will impose now a max weight constraint of 30%

 
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm.res2 = vpsd.res2 = mweights.res2 <- c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns,pm=mutarget,
                  reshigh = rep(0.1,ncol(returns))),silent=TRUE)
  if(class(out)!="try-error"){
    vpm.res2 <- c(vpm.res2,out$pm)
    vpsd.res2 <- c(vpsd.res2,out$ps)
    mweights.res2 <- rbind( mweights.res2,out$pw)
  }
}
# assign the stock names to the column names of mweights
colnames(mweights.res2) <- colnames(returns)
 
##################################################################################
# To avoid the concentration, we will impose now a max weight constraint of 5%


 
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm.res = vpsd.res = mweights.res <- c()

for( mutarget in grid ){
  out = try(portfolio.optim(x = returns,pm=mutarget,
                  reshigh = rep(0.05,ncol(returns))),silent=TRUE)
  if(class(out)!="try-error"){
    vpm.res <- c(vpm.res,out$pm)
    vpsd.res <- c(vpsd.res,out$ps)
    mweights.res <- rbind( mweights.res,out$pw)
  }
}
# assign the stock names to the column names of mweights
colnames(mweights.res) = colnames(returns)


# 
plot(stocksd,stockmu,col="gray" , xlab="Standard deviation (monthly)", ylab="Average return (monthly",
     xlim=c(0.02,0.15), ylim=c(min(stockmu),max(stockmu)))
text( stocksd ,  stockmu ,  labels=colnames(returns), cex= 0.7)
# all portfolios offering less return than the minimum variance portfolio should be excluded
minvarpm <- vpm[vpsd==min(vpsd)]
minvarpm.res <- vpm.res[vpsd.res==min(vpsd.res)]
minvarpm.res2 <- vpm.res2[vpsd.res2==min(vpsd.res2)]
lines(vpsd,vpm,col="black",lwd=1,lty=3)
lines(vpsd.res2,vpm.res2,col="red",lwd=1,lty=3)
lines(vpsd.res,vpm.res,col="blue",lwd=1,lty=3)
lines(vpsd[vpm>=minvarpm],vpm[vpm>=minvarpm],col="black",lwd=2)
lines(vpsd.res2[vpm.res2>=minvarpm.res2],vpm.res2[vpm.res2>=minvarpm.res2],col="red",lwd=2)
lines(vpsd.res[vpm.res>=minvarpm.res],vpm.res[vpm.res>=minvarpm.res],col="blue",lwd=2)

```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```


--- type:NormalExercise lang:r xp:100 skills:1   key:897a83541a
## The minimum variance and maximum Sharpe ratio portfolio

Let us investigate further the previous output consistent of the vector of expected portfolio returns (`vpm`), the 
portfolio standard deviations (`vpsd`) and the matrix of portfolio weights (`mweights`).   

 


*** =instructions
- Complete the sample code to obtain the volatility and expected return of the minimum variance portfolio. 
- Assume a risk free rate of 0.75%. Define `vsr` as the vector holding, for each of the optimized portfolios, the Sharpe ratio computed using `vpm` and `vsd`
- Complete the sample code to compute the risk and reward of the maximum Sharpe ratio portfolio 
- Comple the sample code to compare the allocation of the maximum Sharpe ratio portfolio with the minimum variance portfolio using a barplot (focus hereby on the weights larger than 1%).
 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
stockmu <- colMeans(returns)

grid = seq(min(stockmu),max(stockmu),length.out=50 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <-  c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
colnames(mweights) <- colnames(returns)
```

*** =sample_code
```{r}
# vpm and vd are preloaded
# minimum variance portfolio
pm_minvar <- vpm[vpsd==min(vpsd)]
psd_minvar <-
weights_minvar <-
# Sharpe ratio 
rf <- 0.0075
vsr <- 
# max sharpe ratio portfolio  
pm_maxsharpe <-  
psd_maxsharpe <-
weights_maxsharpe <-
#compare weights
par(mfrow=c(2,1), mar=c(3,2,2,1))
barplot(weights_minvar[weights_minvar>0.01])
 
```

*** =solution
```{r}
######## Josi Concept / potential simplify #######
#combine vectors of expected returns & standarad deviation
grid_portfolios <- data.frame(vpm, vpsd)
colnames(grid_portfolios) <- c("er", "sd")

#compute sharpe ratio 
#sharpe = E[R] / portfolio volatility
#initialze risk free rate of 0.75% first
rf <- .0075
#calculate sharpe ratio
grid_portfolios$sr <- (grid_portfolios$er - rf) / grid_portfolios$sd

#now find minimum variance portfolio
min_var_pf_index <- which(grid_portfolios$sd == min(grid_portfolios$sd))
min_var_pf <- grid_portfolios[min_var_pf_index, ]

#now find max sharpe ratio portfolio 
max_sr_pf_index <- which(grid_portfolios$sr == max(grid_portfolios$sr))
max_sr_pf <- grid_portfolios[max_sr_pf_index, ]

########## END ##########




# solution code
# minimum variance portfolio
pm_minvar <- vpm[vpsd==min(vpsd)]
psd_minvar <-vpsd[vpsd==min(vpsd)]
weights_minvar <- mweights[vpsd==min(vpsd),]
# Sharpe ratio 
rf <- 0.0075
vsr <- (vpm-rf)/vpsd
# max sharpe ratio portfolio  
# max sharpe 
pm_maxsharpe <-  vpm[vsr==max(vsr)]
psd_maxsharpe <- vpsd[vsr==max(vsr)]
weights_maxsharpe <- mweights[vsr==max(vsr),]
# compare weights
par(mfrow=c(2,1), mar=c(3,2,2,1))
barplot(weights_minvar[weights_minvar>0.01])
barplot(weights_maxsharpe[weights_maxsharpe>0.01]) 
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```



--- type:VideoExercise lang:r xp:50 skills:1 key:ca405aabc3c21c142aaae409da6302b77b5210ef
## In-sample versus out-of-sample evaluation


*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v2/hls-ch4_3.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:69f72e3a22916ec2ef924a7f212d354f28022366
## Split-sample evaluation

The function [window](http://www.rdocumentation.org/packages/stats/functions/window), with arguments `x`, `start` 
and `end`, extracts the subperiod between `start` and `end`  from the object `x`. 

*** =instructions
- Split `returns` into the sample `returns.estim` (ending on `2003-12-31`) and `returns.eval` (starting on `2004-01-01`)
- Compute using `returns.estim` the optimal portfolio under a maximum 10% weight constraint, when the target return is the return of the equally weighted portfolio. 
- Compute using `returns.eval` the optimal portfolio under a maximum 10% weight constraint, when the target return is the return of the equally weighted portfolio.
- Make a scatter plot of the two porfolios to see how different the portfolio weights can be when changing the estimation period. Note that if portfolio weights are identical, they should be on the 45 degrees line. 
 


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim <- 
returns.oos <-   
# optimize portfolios
weights.estim <- portfolio.optim(x =  , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x =   ,reshigh = rep(0.1,ncol(returns.estim)))$pw
#scatterplot with 45 degrees lines
 
abline(a=0,b=1,lty=3)
```

*** =solution
```{r}
# solution code

# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim <- window(returns, start = "1991-01-01", end = "2003-12-31")
returns.oos <- window(returns, start ="2004-01-01", end = "2015-12-31")
# optimize portfolios
weights.estim <- portfolio.optim(x = returns.estim , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x = returns.oos ,reshigh = rep(0.1,ncol(returns.estim)))$pw
#
plot(weights.estim,weights.oos)
abline(a=0,b=1,lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6d3bdc717d87317d8c4ccf468836039b35f78440
## Out of sample performance evaluation

How severe is the optimality loss? Let's compare on the out-of-sample period the mean and variance on the optimized portfolio weights using the estimation sample and the out-of-sample evaluation data. 



*** =instructions
- Use the function [Return.portfolio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.portfolio) to compute the portfolio returns of the portfolio with weights `weights.estim` and monthly rebalancing using the `returns.estim` sample. Call the resulting portfolio returns `returns.estim` and use [table.AnnualizedReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/table.AnnualizedReturns) to evaluate the expected performance based on the estimation data.
- Repeat this on the out-of-sample return data, still using `weights.estim`. This is the real performance you may expect from the optimized portfolios. 


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim <- window(returns, start = "1991-01-01", end = "2003-12-31")
returns.oos <- window(returns, start ="2004-01-01", end = "2015-12-31")
# optimize portfolios
weights.estim <- portfolio.optim(x = returns.estim , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x = returns.oos ,reshigh = rep(0.1,ncol(returns.estim)))$pw

```

*** =sample_code
```{r}
# sample code
# weights.estim, returns.estim and returns.oos are preloaded

# portfolio returns on returns.estim
preturns.estim <- Return.portfolio( , ,rebalance_on="months")
 

# portfolio returns on returns.oos
preturns.oos <-  
 
```

*** =solution
```{r}
# solution code
# portfolio returns on returns.estim
preturns.estim <- Return.portfolio(returns.estim,weights.estim,rebalance_on="months")
table.AnnualizedReturns(preturns.estim)
# portfolio returns on returns.oos
preturns.oos <- Return.portfolio(returns.oos,weights.estim,rebalance_on="months")
table.AnnualizedReturns(preturns.oos)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

 
--- type:NormalExercise lang:r xp:100 skills:1   key:9d55676210
## It ain't over
 
You made it! This journey showed you how to compute returns, analyze portfolio performance and optimize portfolios. These are important skills, since whenever you hold or manage more than one asset, you have a portfolio to consider. 

It is now time to compound and get extra return by exploring other useful R packages for portfolio analysis like [PortfolioAnalytics](). Instead of optimizing the portfolio by minimizing the variance under a return targetg constraint, this package allows you optimize virtually all objective functions under all possible constraints. Even if the solution to the problem is not feasible, it will give you the best possible solution using heuristics such as differential evoluation in the R package [DEoptim](). Since I have been involved in both packages, this last exercise is an act of shameless selfpromotion. 

Have fun! 



*** =instructions
- Load the package PortfolioAnalytics and explore the functionality of this package by entering the instruction `?PortfolioAnalytics`. 
- Read my article on portfolio optimization with DEoptim. 
*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(Biobase)
uri <- "https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Ardia~et~al.pdf"
download.file(uri,"deoptim.pdf",method="internal",mode="wb")

```

*** =sample_code
```{r}
# load and explore the package PortfolioAnalytics

# happy reading!
openPDF("deoptim.pdf")
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```



