---
title       : Performance drivers
description : In addition to studying portfolio performance based on the observed portfolio return series, it is relevant to find out how individual (expected) returns, volatilities and correlations interact to determine the total portfolio performance. 

--- type:VideoExercise lang:r xp:50 skills:1  key:bf1fc7dc803e342b4475c296bbebc2bab9b26f32
## The drivers in case of two assets

*** =video_link
//player.vimeo.com/video/108225030

--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:27fc122215
## Driver 1: The assets' individual performance

We have thus seen hat there are three types of drivers for the portfolio performance: the individual performance of the assets in terms of risk and return, the portfolio weight of each of the assets and the correlation between the asset returns.

Let us now first explore further how the portfolio mean return and volatility depends on the mean return and volatility of the assets in which the portfolio is invested. 

To do so, we consider the case of investing with with a monthly frequency in US equities and/or US bonds. The individual asset returns are available in the workspace as `returns_bonds` and `returns_equities`, together with the returns for the so-called 60/40 portfolio that invests every month 60% in equities and 40% in bonds.  

Which of the following statements about the relationship between portfolio performance (investing in both equities and bonds) and the individual assets' performance (investing in only one of them) is false. 


*** =instructions
- A portfolio is less volatile than the sum of its component volatilities. This implies that the standard deviation of the 60/40 portfolio is less than `0.6*sd(returns_equities)+0.4*sd(returns_bonds)`. We call this the subadditivity property of the standard deviation.    
- The portfolio mean return equals the position weighted sum of the component mean returns, that is, `all.equal(mean(returns_6040) , 0.6*mean(returns_equities)+0.4*mean(returns_bonds))`.  
- None of the above statements is correct. 

*** =hint
The proof is in the pudding. Verify the statements directly in the console and you will find  the solution to the problem.

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])

### plot
weights <- c(0.6,0.4)
prices <- cbind(eq_prices,bond_prices)
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_6040 <- Return.portfolio(returns, weights,rebalance_on="months")
# check: all.equal( as.numeric(returns_6040),as.numeric(0.6*returns[,1]+0.4*returns[,2])  )
prices_6040 <- zoo(cumprod(1+returns_6040),order.by=time(returns_6040))
temp <- c( as.numeric(eq_prices), as.numeric(bond_prices) )
ylim <- c( min( temp ) , max(temp) )
plot.zoo(eq_prices , main="Equity-bond portfolios",ylab="", xlab="",ylim=ylim)
lines(bond_prices,col="red")
lines(prices_6040,col="blue")
legend("topleft", legend=c("Equities (ticker: SPY)","Bonds (ticker: AGG)","60/40 Equities-Bonds") , col=c("black","red","blue") , lwd = 1 )

###
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```

--- type:NormalExercise lang:r xp:50 skills:1     key:69c5664fdb
## Driver 2: the choice of porfolio weights

Investors can optimize the choice of weight to obtain the highest risk-adjusted return, as measured by the portfolio Sharpe ratio.

In the special case of investing the total portfolio value in only two assets, there is only one weight to determine, since the weight on the second asset equals one minus the weight on the first asset. 

Let us do this in the case of a portfolio invested US equities and US bonds. We will be using the brute force approach of trying a large number of possible weights and keeping the weight that yields the highest value for hte portfolio Sharpe ratio.   



*** =instructions
- Complete the R script that runs a brute force search to determine the portfolio weights for which the Sharpe ratio is the highest, assuming a zero risk free rate. 
- Plot the value of the Sharpe ratio as a function of the equity weight in the portfolio, and indicate with a verticle line the value of the equities weight that maximizes the portfolio Sharpe ratio. 

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sample_code
```{r}
# returns_equities and returns_bonds are preloaded

# grid defining possible weights equities 
grid <- seq(0,1,0.01)
print(grid);
# initialize vector in which the Sharpe ratios will be saved
vsharpe <- rep(NA,length(grid))
# initialize counter
i <- 1
#loop over the elements in the search grid
for(w in grid){
  # portfolio returns are monthly rebalanced
  preturns <- w*returns_equities+(1-w)*returns_bonds
  # compute the shapre ratio of preturns
  vsharpe[i] <- 
  # update counter  
  i <- 
}

# plot the obtained Sharpe ratios against the equity weights
plot(  ,   ,xlab="Equity weights",ylab="Ann. Sharpe ratio")
abline(v=grid[vsharpe==max(vsharpe)],lty=3)
```


*** =solution
```{r}
# PerformaneAnalytics is pre-loaded
# returns_equities and returns_bonds are preloaded
grid <- seq(0,1,0.01)
vsharpe <- rep(NA,length(grid))
i <- 1
for(w in grid){
  preturns <- w*returns_equities+(1-w)*returns_bonds
  vsharpe[i] <- SharpeRatio.annualized(preturns)
  #vsharpe[i] <- SharpeRatio.annualized(preturns)
  i <- i+1
}
plot(grid,vsharpe,xlab="weights",ylab="Ann. Sharpe ratio")
abline(v=grid[vsharpe==max(vsharpe)],lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1     key:f4f41f6b81
## Driver 3: The correlation between the asset returns

The third driver of portfolio performance is the correlation between the asset returns. Generally speaking, the correlation tells us the tendency of the two asset returns to move together. 

The value of the correlation has important consequences on the portfolio performance in terms of volatility reduction through diversification. In fact, the lower the correlation, the more succesful the portfolio is in terms of partially offsetting large losses in one asset with only a minor loss or even a gain in the other asset. 

In the extreme case of two identical asset returns, the correlation will be one and there is no diversification potential. In the other extreme case where, if one asset return is above average, the other one is almost always below average, the correlation is negative. It is zero, when the asset returns are independent. 

As a simple exercise, suppose you have an equally weighted portfolio of two assets. Their correlation jumps from 0 to 0.5. Then the portfolio variance:    

*** =instructions
- Decreases
- Increases
- Does not change  

*** =hint
Remember, from the video, that the portfolio variance equals  the squared weight of asset 1 times the variance of asset 1, plus the squared weight of asset 2 times the variance of asset 2, plus two times the product of the weights, the volatilities and their correlation. 

*** =pre_exercise_code
```{r}

```


*** =sct
```{r}
# sct code
test_mc(2) 
```



--- type:NormalExercise lang:r xp:100 skills:1 key:b6a986b0e53bd119933fc39df94f21f2bf1e4318
## Interpreting correlation


Measuring is knowing. Let us now compute the correlation between equity returns and bond returns. We will do this first taking all returns in the sample. 

Next, we account for the dynamic nature of correlations, and compute the time series of correlations that we obtain when estimating the correlation over rolling samples. This is a similar analysis as we did for the time-varying performance evaluation in terms of mean return and volatility. 

*** =instructions
- Plot the equity returns (`returns_equities`) against the bond returns (`returns_bonds`) using the function [chart.Scatter](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.Scatter) with the bond returns on the x-asis. Do you see a relation? 
- Compute first the correlation between the variables `returns_equities` and `returns_bonds` using the standard function [cor](). 
- Repeat this using the function [chart.Correlation](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.Correlation), with as argument the multivariate time series `returns`. 
- Commpute the rolling 24-month estimates of the bond-equity correlation using the function [chart.RollingCorrelation](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/chart.RollingCorrelation). 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
prices <- cbind(eq_prices,bond_prices)
colnames(prices) <- c("equities","bonds")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
returns_bonds <- returns$bonds
returns_equities <- returns$equities
```

*** =sample_code
```{r}
# scatter plot

# calculation using cor

# merge returns_equities and returns_bonds into returns

# calculation and visualization using chart.Correlation

# rolling estimates using chart.RollingCorrelation

```

*** =solution
```{r}
# scatter plot
chart.Scatter(returns_equities,  returns_bonds)
# calculation using cor
cor(returns_equities,  returns_bonds)
# merge returns_equities and returns_bonds into returns
returns <- merge(returns_equities,returns_bonds)
# calculation and visualization using chart.Correlation
chart.Correlation(returns)
# rolling estimates using chart.RollingCorrelation
chart.RollingCorrelation(returns_equities, returns_bonds,width=24)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:VideoExercise lang:r xp:50 skills:1 key:a2866e000af4e20577dc0d48d46765e2da1a7d9b
## The general case using matrix notation

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:080c6863fc
## Making a risk-reward scatter diagram

A realistic portfolio is not only invested in US equities and US bonds. It also seeks for diversification by investing in real estate and commodities, among others. We therefore extend our investment opportunity consisting of the US equities ETF (SPY) and US bonds ETF (AGG) to include also a real estate investment trust (VEIEX) and an ETF tracking the GSCI commodities index (GSG). The monthly returns on those investment are available as the variable `returns` in your workspace. 

A good way to visualize the relative attractivenss of investments in terms of risk and rewards is to make a scatter 
plot of the average returns against the portfolio volatilities. 

We need to compute those average returns and volatilities for each asset. In other words, for each return series, and thus for each column in the multivariate return series `returns`, we need to compute the average value and standard deviations.
These calculations are made easy by using the function `apply` with as first argument the return data file, the second argument is the value of `2` indicating  that the calculation should be column-wise, while the third argument is the name of the function that needs to be applied on each column. 

*** =instructions
- Compute the vector of average returns on those four investments using the R command `apply(returns,2,"mean")` and call this `vmeans`.
- Do the same to compute the vector standard deviations and call this `vsd`.
- Complete the R instructions to produce the scatter plot of average returns with respect to standard deviations. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
re_prices <- re_prices/as.numeric(re_prices[1])
comm_prices <-  comm_prices/as.numeric(comm_prices[1])


### plot

prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
returns <- Return.calculate(prices)
returns <- returns[(-1),]
temp <- c( as.numeric(eq_prices), as.numeric(bond_prices),as.numeric(re_prices), as.numeric(comm_prices)   )
ylim <- c( min( temp ) , 1.4*max(temp) )
plot.zoo(eq_prices , main="Performance across asset classes",ylab="", xlab="",ylim=ylim)
lines(bond_prices,col="red")
lines(re_prices,col="blue")
lines(comm_prices,col="purple")
legend("topleft", legend=c("Equities (ticker: SPY)","Bonds (ticker: AGG)", "Real estate (ticker: VEIEX)",
                           "Commodities (ticker: GSG)") , col=c("black","red","blue","purple") , lwd = 1 ,ncol=1)

###
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# vector of means
vmeans <- 
# vector of standard deviations
vsd    <- 
# scatter plot
plot( , vmeans,col="gray",xlim=c(0,0.10))

text( vsd ,  vmeans ,  labels=colnames(returns), cex= 0.7)` 


```


*** =solution
```{r}
# vector of means
vmeans <- apply(returns,2,"mean")
# vector of standard deviations
vsd <- apply(returns,2,"sd")
# scatter plot
plot(vsd,vmeans,col="gray",xlim=c(0,0.10))
text( vsd , vmeans,  labels=colnames(returns), cex= 0.7)
abline(h=0,lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1  key:b7d7806b87
## The covariance matrix

The assets' volatitilities reflect the individual risks of the asset. What matters for the portfolio investors is however the portfolio risk caused by each of the assets, which depends also on the return volatilities.  

Let us compute and analyze the covariance and correlation matrix on the monthly returns of the four asset classes. 

The monthly returns on the investments are available as the variable `returns` in your workspace, together with the previously computed vector of standard deviations `vsd`. 


*** =instructions
- Use `diag(vsd^2)` to print the diagonal matrix with the variances on the diagonal.
- Use `cov(returns)` to compute the covariance matrix;
- Use `cor(returns)`to compute the correlation matrix;
- Verify that the covariance between bond returns and equity returns equals the product between their standard deviations and correlation. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
# vector of means
vmeans <- apply(returns,2,"mean")
# vector of standard deviations
vsd <- as.numeric(apply(returns,2,"sd"))
```

*** =sample_code
```{r}
# covariance matrix with zeros outside the diagonal
diag_cov <-  
print(diag_cov)
# correct covariance matrix
estim_cov <- 
print(estim_cov)
# correlation matrix
estim_cor <- 
print(estim_cor)
# check
all.equal( estim_cov[1,2] ,     )
```


*** =solution
```{r}
# covariance matrix with zeros outside the diagonal
diag_cov <- diag(vsd^2)
print(diag_cov)
# correct covariance matrix
estim_cov <- cov(returns)
print(estim_cov)
# correlation matrix
estim_cor <- cor(returns)
print(estim_cor)
# check
all.equal( estim_cov[1,2] , estim_cor[1,2]*vsd[1]*vsd[2]   )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1   key:21e408b8c65335ed2599fcb8ff64b5c9b352c374
## Matrix-based calculation of portfolio mean and variance

When $w$ is the column-matrix of portfolio weights, $\mu$ the column-matrix of expected returns and $\Sigma$ the 
return covariance matrix. Then the portfolio expected return is $w'\mu$ and the portfolio variance is $w'\Sigma w$.

Let us verify this in R where matrix multiplication requires to use `%*%` instead of `*`. 

*** =instructions
- Use the function [Return.portfolio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.portfolio) to compute the time series of returns for the monthly rebalanced portfolio that is 40% invested in equities, 40% in bonds, 10% in real estate and 10% in commodities. Call this new variable `preturns`. 
- Compute the annualized mean and volatility of the portfolio return series. 
- To replicate this with matrix notation, first complete the code to compute the column-matrix of portfolio weights and
expected returns and the covariance matrix. 
- Now use the matrix products to compute the annualized mean return and volatility.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
eq_prices <- eq_prices/as.numeric(eq_prices[1])
bond_prices <-  bond_prices/as.numeric(bond_prices[1])
re_prices <- re_prices/as.numeric(re_prices[1])
comm_prices <-  comm_prices/as.numeric(comm_prices[1])


### plot

prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
returns <- Return.calculate(prices)
returns <- returns[(-1),]
temp <- c( as.numeric(eq_prices), as.numeric(bond_prices),as.numeric(re_prices), as.numeric(comm_prices)   )
ylim <- c( min( temp ) , 1.4*max(temp) )
plot.zoo(eq_prices , main="Performance across asset classes",ylab="", xlab="",ylim=ylim)
lines(bond_prices,col="red")
lines(re_prices,col="blue")
lines(comm_prices,col="purple")
legend("topleft", legend=c("Equities (ticker: SPY)","Bonds (ticker: AGG)", "Real estate (ticker: VEIEX)",
                           "Commodities (ticker: GSG)") , col=c("black","red","blue","purple") , lwd = 1 ,ncol=1)

###
newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]


```

*** =sample_code
```{r}
# the monthly returns on the four assets are preloaded as  returns
# define vector of weights
weights <-  
# compute portfolio returns
preturns <- Return.portfolio(returns, weights,rebalance_on="months")
# print annualized mean and volatility

# turn weights and mean returns into a column-matrix
weights <- as.matrix(weights)
mu <- as.matrix(apply(returns,2,"mean"))
# sigma is the estimated covariance matrix
sigma <- 
# transsposed matrix
tweights <- t(weights) 
# annualized portfolio mean
12*(tweights%*% )
# annualized volatility
sqrt(12)*sqrt( tweights%*%sigma%*%  )
```

*** =solution
```{r}
# define vector of weights
weights <-   c(0.4,0.4,0.1,0.1)
preturns <- Return.portfolio(returns, weights,rebalance_on="months")
# Annualized mean and volatility
12*mean(preturns)
sqrt(12)*sd(preturns)
# turn weights and mean returns into a column-matrix
weights <- as.matrix(weights)
mu <- as.matrix(apply(returns,2,"mean"))
# sigma is the estimated covariance matrix
sigma <- cov(returns)
# transsposed matrix
tweights <- t(weights)
# transsposed matrix
tweights <- t(weights) 
# annualized portfolio mean
12*(tweights%*%mu )
# annualized volatility
sqrt(12)*sqrt( tweights%*%sigma%*%weights  )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:VideoExercise lang:r xp:50 skills:1   key:ba9b12d382
## The portfolio's risk budget

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:0931e83c3a402b2fb2d78958981928a327428e4d
## Whodunit

compute the portfolio risk budget... barplot? function [StdDev](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/StdDev)

*** =instructions
- Set `vol_budget <- StdDev(returns,portfolio_method="component",weights=c(0.4,0.4,0.1,0.1))` and print this variable;
- Combine the weights and the percentage risk contributions in a table called `weights_percrisk`;
- Make a barplot and note how different the percentage risk contributions are compared to the portfolio weights. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
eq_prices <- get.hist.quote(instrument="SPY",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
bond_prices <- get.hist.quote(instrument="AGG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
re_prices <- get.hist.quote(instrument="VEIEX",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
comm_prices <- get.hist.quote(instrument="GSG",start=as.Date("2006-08-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")

newdates <- seq(as.Date(time(bond_prices ))[2], length=nrow(bond_prices ), by="1 month") - 2
eq_prices <- xts(eq_prices,newdates)
bond_prices <- xts(bond_prices,newdates)
re_prices <- xts(re_prices,newdates)
comm_prices <- xts(comm_prices,newdates)
prices <- cbind(eq_prices,bond_prices,re_prices,comm_prices)
colnames(prices) <- c("equities","bonds","realestate","commodities")
# calculate returns
returns <- Return.calculate(prices)
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code

vol_budget <- StdDev(returns,portfolio_method="component",weights=c(0.4,0.4,0.1,0.1))
out <- cbind( c(0.4,0.4,0.1,0.1) ,  vol_budget$pct_contrib_StdDev)
colnames(out) <- c( "weights" , "perc risk contrib")
barplot(out)
 
  
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

