---
title       : The building blocks 
description : Asset returns and portfolio weights; those are the building blocks of a portfolio return. This chapter is about computing those portfolio weights and returns in R.  

--- type:VideoExercise lang:r xp:50 skills:1 key:b1edf0455cddc24b5691f41bc0bb1b0e9f7e4c1a
## Welcome to the course

*** =video_link
//player.vimeo.com/video/108225030

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:5df9cc8be3e5717ab2d100eb52a85328bc27d243
## Risk and return

Which of the following statements is false:

*** =instructions
- Because most investors do not like risk, risky stocks sell on average at lower market prices.
- Investment risk is time-invariant: it is not influenced by the changes in the business cyle, investor sentiment or other changes in the  market environment. 
- If all assets have the same expected return and risk, an investor is better off by investing in a portfolio with many assets than a portfolio invested in just a few assets. 

*** =hint


*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:NormalExercise lang:r xp:100 skills:1 key:3eba4bd8de6fe4123ea68cc159aa5e8a117fc27d
## Get a feel for the data

The choice of investment matters, even when the underlying risky assets are similar. As a first example, let us consider the stock price of the Coca Cola Company and the PepsiCo company from Januari 2003 till today.  

The time series plot shows you the value evolution of one dollar invested in each of them. As an exercise, plot the time series showing the relative value of an investment in the Coca Cola company, compared to the value of an investment in PepsiCo. To do this exercise, you can us the corresponding price series, available as the variables `ko` and `pep` in your workspace. 

*** =instructions
- Define `ko_pep` as the ratio expressing the value of the share price of the Coca Cola company in terms
of the share price of PepsiCo
- Use `plot.zoo` to visualize the daily variation in this ratio. 
- use `abline` to include a horizontal line at `h=1`.
- Note that, when the value of the ratio increases, `ko` outperforms `pep` and vice versa
- Note also that, When the value of the ratio is larger than one, the total accumulated performance of `ko` since Jan 2003 is higher than that of `pep`. Therefore, as a reference line, use `abline` to include a horizontal line at `h=1`.


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
ko <- get.hist.quote(instrument="KO",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
pep <- get.hist.quote(instrument="PEP",start=as.Date("2003-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)

# 
ko <- ko/as.numeric(ko[1])
pep <- pep/as.numeric(pep[1])

#
plot.zoo(ko, ylim=c( min(c(min(ko),min(pep))) ,  max(c(max(ko),max(pep))) ) , ylab="", xlab="")
lines(pep,col="red")
legend("topleft", legend=c("Coca-Cola","PepsiCo") , col=c("black","red") , lwd = 1 )
```

*** =sample_code
```{r}
# ko and pep are pre-loaded
# Define ko_pep 
ko_pep <- 
# Make a time series plot of ko_pep
  
# Note that, when the value of the ratio increases, the price of the share of CocaCola Company increases more than the price of PepsiCo, and vice versa

# Add, as a reference a horizontal line at 1
  
# When the value of the ratio is larger than one, the cumulative performance of an investment in the CocaCola Company since Jan 2003 is higher than that of an investment in PepsiCo. 
```

*** =solution
```{r}
# Define ko_pep 
ko_pep <- ko/pep
# Make a time series plot of ko_pep
plot.zoo(ko_pep)  
# Note that, when the value of the ratio increases, the price of the share of CocaCola Company increases more than the price of PepsiCo, and vice versa

# Add, as a reference a horizontal line at 1
abline(h=1)  
# When the value of the ratio is larger than one, the cumulative performance of an investment in the CocaCola Company since Jan 2003 is higher than that of an investment in PepsiCo. 
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:d2c7c64108eb91da5a95275fe980d5b4aee3be95
## A first portfolio: buy-and-hold versus daily rebalancing

The choice of investment matters, even when the underlying risky assets are similar. As a second example, let us now consider the stock price of Apple and Microsoft from Januari 2006 till today. The time series plot shows you the value evolution of one dollar invested in each of them. 

It is time to consider our first portfolio. The portfolio approach that we consider is to invest in January 2003, half of the investment budget in Apple and the other half in Microsoft. Since the portfolio weights will change, the investor has two choices. The first one is to be passive and not buy or trade anymore. This is called the a buy and hold investment strategy. Another possibility is to buy and trade at the close of each day in such a way that every day the portfolio is equally invested in the shares of Microsoft and the shares of Apple. We say that this portfolio is daily rebalanced.  

Which of the following statements is false?

*** =instructions
- By investing in a portfolio of risky assets, the portfolio payoff will be in between the minimum and maximum of the payoff of the underlying risky assets.
- By investing in a portfolio of risky assets, the risk of the portfolio payoff will be less than the maximum risk of the underlying risky assets.
- An investor needs to have a lot of capital to be able to invest in a portfolio of many assets. 
- Because the prices of Microsoft and Apple have evolved differently, the investor who spent initially half of his wealth on Microsoft and Apple ends up with a portfolio that is no longer equally invested. 

*** =hint


*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
library(PerformanceAnalytics)
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
aapl <- get.hist.quote(instrument="AAPL",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
msft <- get.hist.quote(instrument="MSFT",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
# 
aapl <- aapl/as.numeric(aapl[1])
msft <- msft/as.numeric(msft[1])
mixed <- 0.5*aapl + 0.5*msft

a <- CalculateReturns(aapl, method="simple")
b <- CalculateReturns(msft, method="simple")
ab <- 0.5*(a+b)
ab[1] <- 0
ew <- cumprod(1+ab)
time(ew) <- time(aapl)
#
plot.zoo(aapl, ylim=c( min(c(min(aapl),min(msft),min(ew),min(ab))) ,  max(c(max(aapl),max(msft),max(ew),max(ab))) ) , ylab="", xlab="")
lines(msft,col="red")
lines(mixed,col="blue")
lines(ew,col="purple")
legend("topleft", legend=c("Apple","Microsoft","50-50 buy and hold","50-50 rebalanced") , col=c("black","red","blue","purple") , lwd = 1 )
```

*** =sct
```{r}
test_mc(3) # if 2 is the correct option.
```


--- type:VideoExercise lang:r xp:50 skills:1 key:8edc78ed1db55bc7450ceaeb586c9c5f4be2ffb2
## The investment decision in a portfolio context

*** =video_link
//player.vimeo.com/video/108225030


--- type:NormalExercise lang:r xp:100 skills:1 key:85a3e5f4f4ca30b9944167e9cbf5eb13692aa7bc
## Calculating portfolio weights when component values are given

An investor has  4000 USD invested in equities, 4000 USD in bonds and 2000 USD in commodities. Compute the weights as the proportion invested in each of those three assets. 

*** =instructions
- Define the vector `values` as the vector holding these three values.
- Print the vector `weights` as the vector `values` divided by the total value (obtained by summing over the component values using the function `sum`.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Define the vector values
values <- 
# Define the vector weights
weights <- 
# Print the resulting weights

```

*** =solution
```{r}
# Define the vector values
values <- c(4000,4000,2000)
# Define the vector weights
weights <- values/sum(values)
# Print the resulting weights
print(weights)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:ca2625a363bea895fda8cc4f1d12f3a2cca25da3
## Defining the weights of an equally weighted portfolio

The equally weighted portfolio is one of the most used portfolios. Suppose the portfolio is equally invested in $N$ assets. Let `N <- 100`. By testing this in the R console, find out which of the following commands does not define properly the weights for the equally-weighted portfolio.  

*** =instructions
- `weights <- rep(1,N)/N`
- `weights <- 1/N`
- `weights <- rep(1/N,N)`

*** =hint
The function `rep(arg1,arg2)` creates a vector in which `arg1` is replicated `arg2` times. 

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ccb1301f85e423d2c1f28ba595bf9e688611fff5
## Defining the weights of a market capitalization weighted portfolio

In a market capitalization weighted portfolio, the weight is given by the asset's market capitalization divided by the sum of the market capitalization of all assets. A typical example is the S&P 500 invested in the 500 largest companies listed on the US stock exchanges (NYSE, Nasdaq).  Note that dividing by the total value of the characteristic across all assets guarantees that the portfolio weights sum up to unity. 

As an exercise, inspect the distribution of market capitalization based weights when the portfolio is invested in 10 stocks and the market capitalizations are 5, 8, 9, 20, 25, 100, 100, 500, 700 and 2000 million USD. 

*** =instructions
- Define the vector `marketcaps` holding the market capitalization
- Define `weights`
- Print `summary(weights)` and make a histogram

*** =hint
A histogram in R can be made using the function [hist()](http://www.rdocumentation.org/packages/graphics/functions/hist).

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Define marketcaps
marketcaps <-  
# Compute the weights
weights <-
# Inspect summary statistics
 
```

*** =solution
```{r}
# Define marketcaps
marketcaps <- c(5, 8, 9, 20, 25, 100, 100, 500, 700, 2000)
# Compute the weights
weights <- marketcaps/sum(marketcaps)
# Inspect summary statistics 
summary(weights)
hist(weights)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```



--- type:VideoExercise lang:r xp:50 skills:1 key:e8205bb01833692ce3cb610117bb029180e68ee8
## The return of a portfolio is the weighted average return
   

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:7a2a60bffa8920ebf2121e0ec548f767197060f0
## Calculation of portfolio returns

As first exercise, we verify that the portfolio return can be computed as the weighted average of the component returns. We do this for the simple case of three assets. 


*** =instructions
- Assume an investment in three assets, for which the initial values are given by `in_values <- c(1000,5000,2000)`.
- Suppose further that the value changes to 1100, 4500 and 3000, respectively. 
- Use the simple return definition to compute the vector of returns on the three component assets
- Use the portfolio return formula to compute the (simple) portfolio return
- Verify that the weighted average portfolio return  equals the percentage change in the total portfolio value.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Vector of initial and final values of the assets
in_values <-  
fin_values <-  
# Vector of simple returns of the assets 
returns <-  
# Weights as the proportion of total value invested in each assets
weights <-
# Compute portfolio return using the portfolio return formula
preturn <- 
# Verify that it matches with the percentage change in the portfolio value
all.equal( preturns   , (sum(fin.values)-sum(in.values))/sum(in.values) )
```

*** =solution
```{r}
# Vector of initial and final values of the assets
in_values <- c(1000,5000,2000)
fin_values <- c(1100,4500,3000)
# Vector of simple returns of the assets 
returns <- (fin_values-in_values)/in_values
# Weights as the proportion of total value invested in each assets
weights <-  in_values/sum(in_values)
# Compute portfolio return using the portfolio return formula
preturn  <- sum(weights*returns)
# Verify that it matches with the percentage change in the portfolio value
all.equal( preturn   , (sum(fin_values)-sum(in_values))/sum(in_values) )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:f40515851bfc2aae1bb2f094ba4f5b8d51c5d272
## Defining the portfolio returns for an equally weighted portfolio

Below you find three commands to compute the portfolio return of an equally weighted portfolio of `N <- 100` assets with return vector `returns`. By testing this in the R console, find out which of the following commands is false.

*** =instructions
- `preturn  <- mean(returns)`
- `preturn  <- sum(rep(1/N,N)*returns)`
- `preturn  <- rep(mean(returns), N)`

*** =hint
The portfolio return is a single number. 

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(1) # if 2 is the correct option.
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:6fda9fa37a19f2326c9e7164c8617f1936dc7124

## From simple to gross returns   

The simple return $R$ expresses the percentage change in value of an investment. The corresponding so-called "gross" return is defined as the future value of $1 invested in the asset for one period, and is thus equal to $1+R$. 

Suppose the investor now has a horizon of two periods and in the first period, she makes a 10% return, but in the second period a loss of 5%. Use the R console to compute the end value of an initial investment of $1000 investment.  Which of the three following answers do you obtain?


*** =instructions
- 1155
- 1045
- 1050

*** =hint
If $R_1$ is the return in the first period and $R_2$ is the return in the second period, then the end-value of a $1 investment is $(1+R_1)*(1+R_2)$.

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:a189cd245c6e378938cab1356e5c01fcebd9b358

## The asymmetric impact of gains and losses  

It is important to be aware of the fact that a positive and negative return of the same magnitude do not compensate each other in terms of terminal wealth. Mathematically, this can be seen from the identity $(1+x)*(1-x)=1-x^2$, which is less than one. A 50% loss is thus not compensated by a 50% gain.  After a loss of 50%, what is the return needed to be at par again? Verify your answer in the R console.

*** =instructions
- 75%
- 100%
- 200%

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```


--- type:VideoExercise lang:r xp:50 skills:1 key:c9e758551130f685dc22840b3ede8cefc4642902
## The PerformanceAnalytics package


*** =video_link
//player.vimeo.com/video/108225030


--- type:NormalExercise lang:r xp:100 skills:1 key:1d4270da1d39343116d0e77f9b897e6b2127914f
## Computing the time series of asset returns  

Computing the returns for one period is thus straightforward to do in R. When the returns need to be calculated for different dates, the functions  [CalculateReturns]() and [Return.portfolio]() in the R package [PerformanceAnalytics](http://www.rdocumentation.org/packages/performanceanalytics/functions/performanceanalytics-package) are helpful. They require the input data to be of the [xts](http://www.rdocumentation.org/packages/xts/functions/xts-package)-time series class.

Let us explore this functionality in the case of a portfolio invested in the Apple stock and Microsoft stock, for which the prices are available in the variables `aapl` and `msft` in your workspace.  

*** =instructions
- Load   the packages `xts` and `PerformanceAnalytics` in your R session.
- Use the function [merge](http://www.rdocumentation.org/packages/xts/functions/merge) to bind the variables `aapl` and `msft` by row names (here: the dates) and call this `prices`.
- Print the first and last six row of `prices` using `head` and `tail`.
- Use the function [CalculateReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.calculate) with as only argument `prices` to compute for each date the return as the percentage change in the price compared to the previous date.
 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)
aapl <- get.hist.quote(instrument="AAPL",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
msft <- get.hist.quote(instrument="MSFT",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
```

*** =sample_code
```{r}
# aapl and msft are pre-loaded

# load the packages xts and PerformanceAnalytics 

# define prices as the merged time series of aapl and msft
prices <-  
# print the first and last six rows of prices
 
  
# define the variable returns holding in each row the returns for that day of Apple and Microsoft
returns <-   
# print the first six rows of returns and note that the first observation is NA, because there is no preceding price needed to compute the return

# remove the first row
returns <- returns[-1,]
 
```

*** =solution
```{r}
# load the packages xts and PerformanceAnalytics 
library(xts)
library(PerformanceAnalytics)
# define prices as the merged time series of aapl and msft
prices <-  merge(aapl,msft)
# print the first and last six rows of prices
returns <-  CalculateReturns(prices)  
# remove the first row
returns <- returns[-1,]
 
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```



--- type:NormalExercise lang:r xp:100 skills:1   key:7fd5556f94
## Computing the time series of portfolio returns (to improve)

Let us now continue the exercise by computing the portfolio returns under different choices of rebalancing the portfolio weights.

The first choice is the buy-and-hold. 

The second choice is to rebalance at a monthly frequency. The target weights are assumed to be the same for each rebalancing date. We call this the constant mix portfolio allocation strategy.   


*** =instructions
- Define `weights` as the vector of equal weights for a portfolio fully invested in two assets. 
- Use the function [Return.portfolio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.portfolio) with as first argument `returns` and as second argument `weights` to compute for each date the return of the buy-and-hold portfolio for which the initial weights are set to `weights`.
- Do the same but for the monthly rebalanced portfolio.
- Make a time series plot of the portfolio returns. 
- Repeat the above, but now set the arguments `verbose=TRUE`. It will now return an object holding various arguments. Investigate this using `names()`
- Compare the end-of period weights
*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)
library(PerformanceAnalytics)
aapl <- get.hist.quote(instrument="AAPL",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
msft <- get.hist.quote(instrument="MSFT",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)

# define prices as the merged time series of aapl and msft
prices <-  merge(aapl,msft)
# print the first and last six rows of prices
returns <-  CalculateReturns(prices)  
# remove the first row
returns <- returns[-1,]
colnames(returns) <- c("AAPL","MSFT")
```

*** =sample_code
```{r}
# returns is pre-loaded
# define the vector of equally-weighted portfolio weights
weights <-  
# compute portfolio return for each date
preturns_bh <-  
preturns_rebal <-  
# plot the time series of portfolio returns
par(mfrow=c(2,1),mar=c(2,4,2,2))
plot.zoo(preturns_bh)
plot.zoo(  )
# verbose=TRUE 
out_bh <-  
out_rebal <-  
# who are you
names(out_bh)
#
endWeights_bh <- out_bh$EOP.Weight 
endWeights_rebal <- out_rebal$EOP.Weight 
# plot the eop weights of Apple
plot.zoo(endWeights_bh$AAPL ,ylim=c(0,1))
plot.zoo(endWeights_rebal$AAPL , ylim=c(0,1) )
```

*** =solution
```{r}
# define the vector of equally-weighted portfolio weights
weights <- c(0.5,0.5)
# compute portfolio return for each date
preturns_bh <- Return.portfolio(returns, weights )
preturns_rebal <- Return.portfolio(returns, weights,rebalance_on="months")
# plot the time series of portfolio returns
par(mfrow=c(2,1),mar=c(2,4,2,2))
plot.zoo(preturns_bh)
plot.zoo(preturns_rebal)
# verbose=TRUE 
out_bh <- Return.portfolio(returns, weights,verbose=TRUE)
out_rebal <- Return.portfolio(returns, weights,rebalance_on="months",verbose=TRUE)
# who are you
names(out_bh)
# eop weights
endWeights_bh <- out_bh$EOP.Weight 
endWeights_rebal <- out_rebal$EOP.Weight 
plot.zoo(endWeights_bh$AAPL ,ylim=c(0,1))
plot.zoo(endWeights_rebal$AAPL , ylim=c(0,1) )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:e2b13d9ade5d7b168bb6157909da1cffbc6729e1
## From daily to monthly portfolio returns

We have now plotted the daily return series. In practice, portfolio managers rarely communicate their daily returns, but focus on the total (compound) performance over a longer period, such as a month or a year. 

The next chapter is devoted to performance evaluation.  But let us  touch intuition now by focussing on the cumulative return evoluation in September 2008. This was an exceptional month for investors, with the bankruptcy of Lehman Brother on September 15 as an important turning point.  

*** =instructions
- Print the daily returns in September 2008 using `preturns["2008-09"]`.
- In order to see how a one dollar investment evolved over that month, print `cumprod(1+preturns["2008-09"])`. Its $k$-th observation  corresponds to the chain $(1+R_1)*(1+R_2)*...*(1+R_K),$ with $R_i$ the $i$-th return of that month
- Compute the return over this month as the simple percentage change between the final and initial value of this one-dollar investment.
- Verify that you obtain the same number as when using the function `Return.cumulative(preturns["2008-09"])]`.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
library(tseries)
aapl <- get.hist.quote(instrument="AAPL",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
msft <- get.hist.quote(instrument="MSFT",start=as.Date("2005-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T)
# load the packages xts and PerformanceAnalytics 
library(xts)
library(PerformanceAnalytics)
# define prices as the merged time series of aapl and msft
prices <-  merge(aapl,msft)
# print the first and last six rows of prices
returns <-  CalculateReturns(prices)  
# remove the first row
returns <- returns[-1,]
# define the vector of equally-weighted portfolio weights
weights <- c(0.5,0.5)
# compute portfolio return for each date
preturns <- Return.portfolio(returns, weights,rebalance_on="days")
```

*** =sample_code
```{r}
# preturns is pre-loaded

# Print the daily returns in September 2008
preturns["2008-09"]
# Print the value evolution of a $1 investment over this month 

# Compute the monthly return as the percentage change of final value w.r.t. initial value

# Verify that this leads to the same results as using the function Return.cumulative

```

*** =solution
```{r}
# Print the daily returns in September 2008
preturns["2008-09"]
# Print the value evolution of a $1 investment over this month 
cumprod(1+preturns["2008-09"])
# Compute the monthly return as the percentage change of final value w.r.t. initial value
(0.7682575-1)/1
# Verify that this leads to the same results as using the function Return.cumulative
Return.cumulative(preturns["2008-09"])
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:9953e9ca351425f91036a223de3ba9ee04c5a282
## Sneak preview of the portfolio fin and fun ahead

The *fin*ancial theory of portfolio analysis and the *fun* of applying portfolio management in R, that's what this course is all about.

It's about the landmark contributions of Nobel prize winners like Daniel Kahneman, Harry Markowitz, William Sharpe, James Tobin and Robert Engle. 

To maky you hungry for more, look at the nice output tables and plots printed in the workspace. This is what we will be doing in the next chapter, leveragin mainly the functionality in the package [PerformanceAnalytics](). 

Which of the following functions is not in the package [PerformanceAnalytics](). 


*** =instructions
- According to Daniel Kahnemen, in decision involving risk, most people are anything but rational. They fail to have the broad view and are carried away by individual gains and losses. You win a few, you lose a few. It's the bottom line that matters. This is to be computed using: [Return.annualized](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.annualized). 
- [SharpeRatio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/SharpeRatio).  Shapre and tobin. 
- [optim.portfolio](http://www.rdocumentation.org/packages/tseries/functions/optim.portfolio]). Harry Markowitz of the portfolio returns compared to the returns of a benchmark strategy, the higher the risk that the returns on the portfolio strategy will deviate from the benchmark returns. 
- Rolling standard deviations. Engle  

*** =hint
hint

*** =pre_exercise_code
```{r}
library(PerformanceAnalytics)
```

*** =sct
```{r}
test_mc(1) 
```
 






