---
title       : The building blocks 
description : Asset returns and portfolio weights; those are the building blocks of a portfolio return. In this chapter, you will first learn about computing portfolio weights in R. Next we will see how to combine portfolio weights and asset returns into the calculation of portfolio returns. The exercises involve both simple two-asset cases and a real-life portfolio invested in the 30 Dow Jones Industrial Average stocks. The final exercises consist in linking portfolio returns to the nominal portfolio value and doing a multi-period analysis.   

--- type:VideoExercise lang:r xp:50 skills:1
## The power of portfolio diversification

*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1
## Calculating portfolio weights when component values are given

An investor has  4000 USD invested in equities, 4000 USD in bonds and 2000 USD in commodities. In order to compute the corresponding portfolio weights, define the vector `values` as the vector holding these three values. Then define the vector `weights` as the vector `values` divided by the total value (obtained by summing over the component values using the function `sum`).

*** =instructions
- Define the vector `values`.
- Define the vector `weights`.
- Print the vector `weights`.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Define the vector values
values <- 
# Define the vector weights
weights <- 
# Print the resulting weights
print(weights)
```

*** =solution
```{r}
# Define the vector values
values <- c(4000,2000,2000)
# Define the vector weights
weights <- values/sum(values)
# Print the resulting weights
print(weights)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Defining the weights of an equally weighted portfolio

The equally weighted portfolio is one of the most used portfolios. Suppose the portfolio is equally invested in $N$ assets. Let `N <- 100`. Which of the following commands does not define properly the weights for the equally-weighted portfolio.  

*** =instructions
- `weights <- rep(1,N)/N`
- `weights <- 1/N`
- `weights <- rep(1/N,N)`

*** =hint
The function `rep(arg1,arg2)` creates a vector in which `arg1` is replicated `arg2` times. 

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:NormalExercise lang:r xp:100 skills:1
## Defining the weights of a market capitalization weighted portfolio

In a market capitalization weighted portfolio, the weight is given by the asset's market capitalization divided by the sum of the market capitalization of all assets. A typical example is the S&P 500 invested in the 500 largest companies listed on the US stock exchanges (NYSE, Nasdaq).  Note that dividing by the total value of the characteristic across all assets guarantees that the portfolio weights sum up to unity. 

As an exercise, inspect the distribution of market capitalization based weights when the portfolio is invested in 10 stocks and the market capitalizations are 5, 8, 9, 20, 25, 100, 100, 500, 700 and 2000 million USD. 

ofdefine the vector `marketcaps` as the sequence from 1 to 500 and compute the corresponding market capitalization based portfolio weights. Inspect the distribution of the portfolio weights using the function `summary()`.

*** =instructions
- Define the vector `marketcaps` holding the market capitalization
- Define `weights`
- Print `summary(weights)` and make a histogram

*** =hint
A histogram in R can be made usin the function `hist`.

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Define marketcaps
marketcaps <- c(5, 8, 9, 20, 25, 100, 100, 500, 700, 2000)
# Compute the weights
weights <-
# Inspect summary statistics
summary(weights)
```

*** =solution
```{r}
# Define marketcaps
marketcaps <- c(5, 8, 9, 20, 25, 100, 100, 500, 700, 2000)
# Compute the weights
weights <- marketcaps/sum(marketcaps)
# Inspect summary statistics
summary(weights)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:VideoExercise lang:r xp:50 skills:1
## Portfolio returns are the weighted average of the individual returns


*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1
## Calculation of portfolio returns

As first exercise, we verify that the portfolio return can be computed as the weighted average of the component returns. We do this for the simple case of three assets and assume the porfolio weights to be value-weighted. 

Assume further that the initial vector of component values is given by `in.values <- c(1000,5000,2000)`. We further suppose that the value changes to `fin.values <- c(1100,4500,3000)`. Define the value weighted portfolio weights as `weights.vw <- in.values/sum(in.values)` compute the component returns using the formula of simple returns. Verify that the weighted average portfolio return ( `sum(in.weights*returns)`  ) equals the percentage change in the total portfolio value ( `sum(fin.values)-sum(in.values))/sum(in.values)` ).

*** =instructions
- Use the simple return definition to compute the vector of returns on the three component assets
- Use the portfolio return formula to compute the (simple) portfolio return

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Initial and final values of the assets
in.values <- c(1000,5000,2000)
fin.values <- c(1100,4500,3000)
# Compute simple returns 
returns <- (fin.values-in.values)/in.values
# Compute value weighed portolio weights
weights.vw <- in.values/sum(in.values)
# Compute portfolio return using the portfolio return formula
preturns.vw <- 
# Verify that it matches with the percentage change in the portfolio value
( preturns.vw == (sum(fin.values)-sum(in.values))/sum(in.values) )
```

*** =solution
```{r}
# Initial and final values of the assets
in.values <- c(1000,5000,2000)
fin.values <- c(1100,4500,3000)
# Compute simple returns 
returns <- 
# Compute value weighed portolio weights
weights.vw <- in.values/sum(in.values)
# Compute portfolio return using the portfolio return formula
preturns.vw <- sum(weights.vw*returns)
# Verify that it matches with the percentage change in the portfolio value
( preturns.vw == (sum(fin.values)-sum(in.values))/sum(in.values) )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Defining the portfolio returns for an equally weighted portfolio

Below you find three commands to compute the portfolio return of an equally weighted portfolio of `N <- 100` assets with return vector `returns`. Which command is wrong? 

*** =instructions
- `preturn.ew <- mean( returns )`
- `preturn.ew <- sum( rep(1/N,N)*returns )`
- `preturn.ew <- rep( mean(returns),N )`

*** =hint
The portfolio return is a single number. 

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(1) # if 2 is the correct option.
```


--- type:NormalExercise lang:r xp:100 skills:1
## Multiperiod returns

Suppose an investor realizes monthly returns of 2.5%, 3%,-2.1%,0.5%,1%,-0.4%,2%,1%,4%,-5%,3%,2%. How much is the annual return? 

*** =instructions
- Compute the annual return by compounding the monthly returns


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}

```

*** =solution
```{r}
(1+0.025)*(1+0.03)*(1-0.021)*(1+0.005)*(1+0.01)*(1-0.004)*(1+0.02)*(1+0.01)*(1+0.04)*(1-0.05)*(1+0.03)*(1+0.02)-1
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:VideoExercise lang:r xp:50 skills:1
## A realistic portfolio case: Investing in the 30 DJIA stocks over 25 years with monthly rebalancing


*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1
## The packages xts and PerformanceAnalytics
Up to now, we have done simple calculations in R. In the future we will need more complex functionality, which is available in the R packages `xts` (for time series analysis in general) and `PerformanceAnalytics` (for portfolio analysis in particular). Load this two packages in your R session.


*** =instructions
- Load the packages


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# Load packages

```

*** =solution
```{r}
# Load packages
library(xts)
library(PerformanceAnalytics)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1
## The multivariate time series of prices and returns 
Let us now inspect the The xts file `prices` containing the end-of-month prices for the 30 Dow Jones Industrial Average constitutuents. 
Make use of the `head(prices,1)` and `tail(prices,1)` functions to print the first and last rows of `prices` and verify that the price data runs from
December 1990 to December 2015. Then compute the monthly returns using the function `returns <- Return.calculate(prices, method="simple")`. Print the first 6 rows of returns and verify that the first row consists of NA's. Remove the first row from returns using `returns <- returns[(-1),]`. 


*** =instructions
- Print the first and last row using the `head` and `tail` functionality 
- Define `returns` as the multivariate return series using  `Return.calculate` with the option of simple returns
- Remove the first observation in `returns`

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds")
```

*** =sample_code
```{r}
# First and last row
head(prices,1)
tail(prices,1)
# Compute the simple returns based on the prices of DJIA stocks
returns <- 
# Remove the first row
returns <- 
```

*** =solution
```{r}
# First and last row
head(prices,1)
tail(prices,1)
# Compute the simple returns based on the prices of DJIA stocks
returns <- CalculateReturns(prices, method="simple")
# Remove the first row
returns <-returns[(-1),]
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1
## Plotting and tabulating the time series of portfolio returns

Let us now consider the portfolio that invests equal amounts in each of the 30 stocks and the portfolio is rebalanced every month. The monthly portfolio return is therefore simply the average return over the month. It can be computed in R using the function `Return.portfolio`, which by default assumes equal investment
weights for each period. This leads to time series of monthly portfolio returns. An effective way to tabule monthly data over saveral years is to use the function `table.CalandarReturns()` in PerformanceAnalytics to present it in the format year x month.  Note the large portfolio losses during the burst of the internet bubble in 2002 and the financial crisis of 2009.  Finally, plot the resulting time series using the function `plot.zoo()`. Note that the time series is centered around zero and that the magnitude of the variations in return changes over time, because of time-varying volatility. 

*** =instructions
- define `preturns.ew` as the time series of monthly returns on equally weighted portfolio of DJIA constituents
- make a month x year table of the `preturns.ew`
- use `plot.zoo()` to make a time series plot of `preturns.ew`

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# Define preturns.ew
preturns.ew <- 

# Tabulate those returns as year x month  

# Make the time series plot

```

*** =solution
```{r}
# Define preturns.ew
preturns.ew <- Return.portfolio( returns )
# Tabulate those returns as year x month  
table.CalendarReturns(preturns.ew)  
# Make the time series plot
plot.zoo(preturns.ew)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1
## The evolution of portfolio value - The booms and busts of the stock market
The function `table.CalendarReturns` already reported in the last column the cumulative return performance per year. In order to obtain the cumulative investment value over a years, one needs to multiply the initial value value with $(1+R_1)*(1+R_2)*(1+R_3)*\ldots*(1+R_{12}),$ where $R_i$ is the return in month $i$. In R such a cumulative product is easily obtained using the function `cumprod(1+returns)` with returns the series of returns that needs to be compounded. Suppose the investor has $1 invested in the strategy of equally weighting all DJIA assets and monthly rebalancing. Plot the time series evolution of that investment over the full period available in the `preturns.ew` time series vectors. 


*** =instructions
- Define `cumulative.value` as the cumulative product of the total returns 
- Plot this wealth evolution using the function `plot.zoo' and note the drawdowns during the market crashes.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds")
returns <- CalculateReturns(prices, method="simple")
preturns.ew <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# Define the time series holding the cumulative.value
cumulative.value <-    
# plot the wealth evolution

```

*** =solution
```{r}
# Define the time series holding the cumulative.value
cumulative.value <- cumprod(1+preturns.ew)
# plot the wealth evolution
plot.zoo(cumulative.value)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## The asymmetric impact of gains and losses

It is important to be aware of the fact that a positive and negative return of the same magnitude do not compensate each other. Mathematically, this can be seen from the identity $(1+x)*(1-x)=1-x^2$, which is less than one. A 50% loss is thus not compensated by a 50% gain.  After a loss of 50%, what is the return needed to be at par again? 

*** =instructions
- 75%
- 100%
- 200%

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```





