---
title       : Interpreting portfolio returns 
description : The history of portfolio returns contains valuable information on how much the investor can expect to gain or lose. Investors summarize these numbers graphically in terms of distribution plots and through statics expressing average return, risk and risk-adjusted return over rolling estimation samples. This chapter concludes by an investigation in the non-normality of the return series of investments and provides tools to characterize the financial risks and rewards that that takes this non-normality into account. 

--- type:VideoExercise lang:r xp:50 skills:1 key:c4b994ff53
## The different dimensions of portfolio performance


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:04255405f1b95fc9587c9350f29c6549b41cc79e
## Balancing risk and reward 

Here you see three return distributions. Note that the distributions are symmetric and the center of the distribution thus corresponds to the average return. Return distributions A and B have the same expected return, but for distribution B the volatility is higher, and therefore the range of possible outcomes is much wider. This indicates that the risk of a large deviation from the mean return is higher for investing in B than when investing in A. Distribution C has the same level of volatility as distribution A but a higher return. 

Consider the typical investor who likes returns, but dislikes risk. Which distribution does she prefer? 


*** =instructions
- Distribution A.
- Distribution B.
- Distribution C.

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
par(mfrow=c(3,1),mar=c(3,3,1,1),cex.axis=1.2)
vd_A <- dnorm(vx,mean=0,sd=0.10)
vd_B <- dnorm(vx,mean=0,sd=0.20)
vd_C <-dnorm(vx,mean=0.10,sd=0.10)
ylims = c(0,max(c(vd_A,vd_B,vd_C)))
vx = seq(-0.5,0.5,0.001)
plot( vx , vd_A,type="l",ylim=ylims,ylab="")
abline(v=0,lty=3)
text(x=-0.4,y=3,labels="Investment A")
#arrows(x0=-2*0.1,y0=dnorm(-2*0.1),x1=2*0.1,y1=dnorm( 2*0.1),lty=3)
#arrows(x0=2*0.1,y0=dnorm(-2*0.1),x1=-2*0.1,y1=dnorm( 2*0.1),lty=3)
plot( vx , vd_B,type="l",ylim=ylims,ylab="")
abline(v=0,lty=3)
text(x=-0.4,y=3,labels="Investment B")
plot( vx , vd_C,type="l",ylim=ylims,ylab="")
text(x=-0.4,y=3,labels="Investment C")
abline(v=0.10,lty=3)
```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```


--- type:NormalExercise lang:r xp:100 skills:1 key:07d324a8e6

## Exploring the monthly S&P 500 returns 

Let us now examine the monthly performance of the S&P 500 portfolio. 

A picture is worth a thousand word. This is why most analyses of performance start by analyzing the time series plot of the value of the investment. For the S&P 500 this is shown in the upper plot for the period 1950 to current data. Each observation is the end-of-day value. Note the booms and busts in the value of this portfolio. Do you see why the 2000s are often called the lost decade of investing?

These prices are available in the R console as the variable `sp500`, which is of the xts time series class.  
Our interest is in describing the monthly performance. This exercise is about aggregation daily price series to end-of-month prices and then computing monthly returns and visualize them in an attractive table.  

*** =instructions
- Define the variable `sp500_monthly` as the output of applying the function [to.monthly](http://www.rdocumentation.org/packages/xts/functions/to.period) to `sp500`.
- Print the first six rows of `sp500_monthly` and note that the aggregation has resulted in a table of four columns, holding the
opening, lowest, highest and close price of `sp500` for each month. 
- Define the variable `sp500_returns` as the close-to-close monthly returns. 
- Make the time series plot of the monthly returns.
- Use the function [table.CalandarReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/table.CalandarReturns) in PerformanceAnalytics to present the monthly return data in the visually attractive format year x month. 
*** =hint


*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="d")
library(PerformanceAnalytics)
sp500 <- xts(sp500)
#par(mfrow=c(2,1),mar=c(3,2,2,2),cex.axis=0.7)
plot.zoo(sp500 , main="Value of the S&P 500 portfolio",ylab="", xlab="")
#plot.zoo(sp500_returns , main="Monthly returns of the S&P 500 portfolio",ylab="", xlab="",type="h")
```

*** =sample_code
```{r}
# Time aggregation

# Print first six rows

# Return calculation

# Time series plot

# Produce the year x month table

```

*** =solution
```{r}
# Time aggregation
sp500_monthly = to.monthly(sp500)
# Print first six rows
head(sp500_monthly)
# Return calculation
sp500_returns <-  xts(CalculateReturns(sp500_monthly[,4]))  
# Time series plot
plot.zoo(sp500_returns)
# Produce the year x month table
table.CalendarReturns(sp500_returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:8c05172eb2

## The monthly mean and volatility  

In this exercise, you need to do a descriptive analysis of the mean and volatility of the monthly S&P 500 returns. These returns are available in the R console as the variable `sp500_returns`. 

We will also compare the returns with

*** =instructions
- Compute the mean value of the return series.
- Compute the standard deviation of the return series. 
- Compute the risk-adjusted return given by the mean return divided by the standard deviation.
*** =hint
The functions `mean` and `sd` are standard functions in R to estimate the mean and standard deviation.

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
library(PerformanceAnalytics)
sp500_returns <-  xts(CalculateReturns(sp500))[-1]
names(sp500_returns) <- "sp500"
```

*** =sample_code
```{r}
# Compute the mean monthly returns 

# Compute the standard deviation
  
# Compute the mean return divided by the return standard deviation

```

*** =solution
```{r}
# Compute the mean monthly returns 
mean(sp500_returns)
# Compute the standard deviation
sd(sp500_returns)  
# Compute the mean return divided by the return standard deviation
mean(sp500_returns)/sd(sp500_returns)
# Produce the year x month table
table.CalendarReturns(sp500_returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

 
--- type:NormalExercise lang:r xp:100 skills:1 key:c14e2a1a3a
 
## Excess returns and the portfolio's Sharpe ratio  

Average returns in itself are difficult to interpret. Is an average of 5% a high return? It depends on the alternatives. Suppose e.g. that there is a risk free investment offerening 4%. Then, by taking risks, the investor only has a 1 percentage point higher average return. The solution is to consider returns in excess of the risk-free investment. Those are called the excess return. They can be interpreted as the return on the self-financed investment of buying the risky portfolio and borrowing at the risk free rate.  

The Sharpe ratio summarizes the balance between the excess portfolio return and the risk taken. It is the ratio obtained by taking the average difference between the portfolio return and the return the investor would have from a risk free investment and dividing this number with the portfolio standard deviation.The Sharpe ratio thus shows the reward per unit of risk taken.  

A good proxy for the risk free rate is the yield of a 1-month Treasury Bill. This Tbill rate is available in the console as the variable `rf` (TO DO: make time series).  The S&P 500 portfolio returns are still available as `sp500_returns`.  
The Sharpe ratio of a portfolio When we take the average excess portfolio return and divide it by the portfolio volatility



*** =instructions
- Compute the series of excess portfolio return `sp500_excess`.
- Compute the Sharpe ratio
*** =hint
The functions `mean` and `sd` are standard functions in R to estimate the mean and standard deviation.

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
library(PerformanceAnalytics)
sp500_returns <-  xts(CalculateReturns(sp500))[-1]
names(sp500_returns) <- "sp500"
rf <- 0.01/12
```

*** =sample_code
```{r}
# Compute the series of excess portfolio return `sp500_excess`.

# Compute the Sharpe ratio
  

```

*** =solution
```{r}
# Compute the series of excess portfolio return `sp500_excess`.

# Compute the Sharpe ratio
  
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1

## Annualized mean and volatility

The mean and volatility of monthly returns corresponds to the average and standard deviation over a monthly investment horizon. Investors annualize those statistics by mutiplying the mean with 12, and the volatility with the square root of twelve. Compute the annualized mean and volatility of the portfolios.

*** =instructions
- Print the annualized mean of `sp500_returns`.
- Print the annualized standard deviation of `sp500_returns`.
- Assume the risk free rate is 0. What is the portfolio's Sharpe ratio?
- Verify that the functions [Return.annualized](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.annualized), [StdDev.annualized](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/StdDev.annualized)  and [table.AnnualizedReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/table.AnnualizedReturns) report the same numbers.
*** =hint
hint comes here

*** =pre_exercise_code
```{r}
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
library(PerformanceAnalytics)
sp500_returns <-  xts(CalculateReturns(sp500))[-1]
names(sp500_returns) <- "sp500"
```

*** =sample_code
```{r}
# Annualized mean

# Annualized standard deviation

# Same answer?

```

*** =solution
```{r}
# Annualized mean
12*mean(sp500_returns)
# Annualized standard deviation
sqrt(12)*sd(sp500_returns)
# Same answer?
Return.annualized(sp500_returns)
StdDev.annualized(sp500_returns)
table.AnnualizedReturns(sp500_returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:8dcb1217fb2989d6c7e3615464b65c0aeae11235
## The annualized Sharpe ratio

The Sharpe ratio is the average excess return per unit of portfolio volatility. Suppose the average excess return and portfolio volatility are computed on monthly returns. What is the effect of using annualized mean excess returns and volatility on the Sharpe ratio? 

*** =instructions
- The annualized Sharpe ratio and the monthly Sharpe ratio are identical.
- The annualized Sharpe ratio is 12 times the monhtly Sharpe ratio.
- The annualized Sharpe ratio is $\sqrt{12}$ times the monhtly Sharpe ratio.

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec

```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```


--- type:NormalExercise lang:r xp:100 skills:1 key:9d2eb4bc6c13e8cdabe71c739978050d8809058c
## Don't be fooled by randomness

In all cases, we aim at inferring from the past returns relevant information on performance. It is important to remember that when doing this we expose ourselves to estimation errors. We will never be able to quantify true performance, we can only make a guess about it from the past returns.  This is the limitations of any estimation approach. It is almost sure that the estimates of the mean and volatility will be different from the true mean and standard deviation.

Familiarize now yourself with the uncertainty in estimation by drawing random samples from a normal distribution with mean of 5% and standard deviation of 10%, and how this uncertainty is reduced when the sample size grows.  

*** =instructions
-  Set the variable `returns` to the sample of 12 observations obtained by using the function `rnorm` to sample from the normal  distribution with mean 5% and standard deviation of 10%.  
- Compute the mean and standard deviation of the return series. 
- Repeat the first two instructions and see how different the numbers are for a different sample. 
- Repeat the first three instructions but for a sample of 60 observations.  
*** =hint


*** =pre_exercise_code
```{r}
# pec

```

*** =sample_code
```{r}
# Draw the random sample

# Compute the mean and standard deviation
  
# Repeat the above

# Repeat the above for a sample of 1000 observations

```

*** =solution
```{r}
# Draw the random sample
returns <- rnorm(12,mean=0.05,sd=0.10)
# Compute the mean and standard deviation
mean(returns)
sd(returns)
# Repeat the above
returns <- rnorm(12,mean=0.05,sd=0.10)
mean(returns)
sd(returns)
# Repeat the above for a sample of 1000 observations
returns <- rnorm(60,mean=0.05,sd=0.10)
mean(returns)
sd(returns)
returns <- rnorm(60,mean=0.05,sd=0.10)
mean(returns)
sd(returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:a481e8c772e346eb869ef192f4b17070be61961e
## Time-variation in portfolio performance


*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:737157d7421c452d9ab7cb48fc219403c4db82ff
## Rolling annualized mean and volatility

charts.RollingPerformance on the portfolio return vector

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# sample code
chart.RollingPerformance(R=preturns , width=12,
                         FUN="Return.annualized",scale=12)
chart.RollingPerformance(R=preturns , width=12,
                         FUN="sd.annualized",scale=12)
```

*** =solution
```{r}
# solution code
chart.RollingPerformance(R=preturns , width=12,
                         FUN="Return.annualized",scale=12)
chart.RollingPerformance(R=preturns , width=12,
                         FUN="sd.annualized",scale=12)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:2169b00c723410a1ee9217836b284bf4cdd7dc2c
## Effect of window length choice

I show two plots, one with rolling windows of 6 months and one with rolling windows of 12 months. Which one is more persistent?

*** =instructions
- Annualized volatility on rolling window of 6 months
- Annualized volatility on rolling window of 12 months

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
library(PerformanceAnalytics)
sp500_returns <-  xts(CalculateReturns(sp500))[-1]
names(sp500_returns) <- "sp500"
par(mfrow = c(2,1))
chart.RollingPerformance(R=sp500_returns , width=6,
                         FUN="sd.annualized",scale=12,main="6-months rolling annualized volatility S&P 500")
chart.RollingPerformance(R=sp500_returns , width=12,
                         FUN="sd.annualized",scale=12,main="12-months rolling annualized volatility S&P 500")
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```

--- type:NormalExercise lang:r xp:100 skills:1 key:012c07fe20924387617cef2df90e9ac6458f6fa8
## Subperiod performance analysis and the function window 

Let us now consider the time-variation in the return distribution. Define `sp500_returns_2008` as the S&P 500 portfolio returns in 2008, obtained with the function `window`, having as first argument the return series that needs to be subset, then the start data and end date in the format `YYYY-MM-DD`. Compare the histogram of `sp500_returns` with `sp500_returns_2008` and well as the annualized volatility. 


*** =instructions
-  Define `preturns.2008`
-  Plot the histograms
-  Commpute annualized volatility

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
options(warn=-1)
library(tseries)# Its function get.hist.quote allows to download prices from Yahoo!Finance
library(xts) # Its function plot.zoo make simple, but attractive, time series plots
# download adjusted close prices (that is corrected for dividend payments and stock splits)
# dates have the format "YYYY-MM-DD"
sp500 <- get.hist.quote(instrument="^GSPC",start=as.Date("1950-12-31"),end=Sys.Date(),quote="AdjClose",quiet=T,compression="m")
library(PerformanceAnalytics)
sp500_returns <-  xts(CalculateReturns(sp500))[-1]
names(sp500_returns) <- "sp500"
```

*** =sample_code
```{r}
# Define preturns.2008
sp500_returns_2008 <- window(sp500_returns,startdate = as.Date(" "), enddate = as.Date(" ")  )
# plotting
par( mfrow = c(2,1) , mar=c(3,2,2,2))

# annualized volatility

```

*** =solution
```{r}
# solution code
# Define preturns2008
sp500_returns_2008 <- window(sp500_returns,startdate = as.Date("2008-01-01"), enddate = as.Date("2008-12-31 ")  )
# plotting
par( mfrow = c(2,1) , mar=c(3,2,2,2))
hist(sp500_returns_2008)
hist(preturns)
# annualized volatility
sqrt(12)*sd(sp500_returns_2008)
sqrt(12)*sd(sp500_returns)         

```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:c694fec6bc34ed0db1ccf2eea79660404a3a1e55
## Analyzing the portfolio return distribution

The variable `preturns` contains the monthly returns on the equally-weighted portfolio of Dow Jones Industrial Average stocks over the period 1991-2015. Compute first the mean and standard deviation of these returns, as well as their minimum and maximum value. Then plot the histogram of these returns.

*** =instructions
- Compute summary statistics using the functions `mean`, `sd`, `min` and `max`
- Plot the histogram and expriment with the argument `breaks` in `hist` to make the intervals smaller

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# Summary statistics




# Histogram


```

*** =solution
```{r}
# Summary statistics
mean(preturns)
sd(preturns)
min(preturns)
max(preturns)
# Histogram
hist(preturns,breaks=10)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```





--- type:VideoExercise lang:r xp:50 skills:1 key:1c105a499d956a288a91e9884d2f1765c20a025f
## Interpreting the non-normality of the return distribution 


*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:dfd18e3152bb168fd30148d4b418df9383e2fa19
## Skewness and kurtosis

Exercise on computing skewness, kurtosis.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:774962285dd78c6a424e6a1eebc8c1cea5949cd3
## Downside risk measures

The use of the VaR function to compute value at risk.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:20cda1434c7729ec212810fc6317b96b99c1b0ec
## Drawdowns

The use of the [table.Drawdowns]() function in PerformanceAnalytics.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:d3d74e3f4e43c9444ca582ded3a6d7f2dc8746e1
## The effect of leverage

Suppose that by leveraging the portfolio, the returns are multiplied by 2. Verify that the mean and standard deviation of the leveraged returns is twice the mean and standard deviation of the original returns and that the skewness and kurtosis are not affected by the levarage. 

*** =instructions
- Define the variable `lpreturns`
- Compute for `preturns` and `lpereturns` the vector holding the mean, standard deviation, skewness and kurtosis



*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
preturns <- Return.portfolio( returns )
```

*** =sample_code
```{r}
# Define lpreturns

# Vector with mean, sd, skewness and excess kurtosis for preturns
c(  ,   ,   ,  )
# Vector with mean, sd, skewness and excess kurtosis for lpreturns
c(  ,   ,   ,  )
```

*** =solution
```{r}
# Define lpreturns
lpreturns = 2*preturns
# Vector with mean, sd, skewness and excess kurtosis for preturns
c(  mean(preturns) , sd(preturns)   , skewness(preturns)  , kurtosis(preturns)  )
# Vector with mean, sd, skewness and kurtosis for lpreturns
c(  mean(lpreturns) , sd(lpreturns)   , skewness(lpreturns)  , kurtosis(lpreturns)  )
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


