---
title       : Optimizing the portfolio 
description : We have up to now considered the portfolio weights as given. In this chapter you learn how to determine in R the portfolio weights that are optimal in terms of achieving a target return with minimum risk, while satisfying constraints on the portfolio weights.  
 
--- type:VideoExercise lang:r xp:50 skills:1 key:2d45c303bc2e49767a9cd744f15a31938a2ec7d9
## Modern portfolio theory of Harry Markowitz


*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v1/hls-ch4_1.master.m3u8




--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:139d7eabbd827860709b29f2a8d968614d1e3cb0
## Exploring the problem of mean-variance based investing in DJIA stocks 

The theory of portfolio optimization will be illustrated on the universe of the 30 Dow Jones Industrial Average stocks. 

Like the S&P 500 portfolio studied in Chapter 2, the Dow Jones Industrial Average portfolio is an important reference portfolio for the US equity market. It is invested in 30 large publicly owned companies based in the United States.

We will be using the sample of monthly returns from January 1991 till December 2015. These return data are available in the workspace as the variable `returns`.   

Assume an investor who only cares about the mean and standard deviation of the portfolio return distribution. She likes expected returns and dislikes volatility.  Which of the following statements is false:

*** =instructions
- If the investor has to choose between investing all her wealth in Home Depot (ticker: HD) versus investing in Caterpillar (ticker: CAT), she will choose to invest in Home Depot, since it has a higher average return and lower risk.  
- The investor will never combine investments in Home Depot and Caterpillar in her portfolio. 
- By combining investments, it may be possible to obtain a portfolio with a lower variance than the variance of the Exxon Mobil Corporation's stock (ticker: XOM). 
- If an investor is not allowed to borrow, then there is no portfolio invested in thirty Dow Jones Industrial Average stocks, offering a higher average return than the one of Apple (ticker: AAPL).   
- There is a total of 300 monthly observations in our sample of monthly returns. The highest return is around 88% while the lowest return is around -58%. 

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

stockmu = apply( X = returns , MARGIN = 2 , FUN="mean")
stocksd = apply( X = returns , MARGIN = 2 , FUN="sd")

# make the scatter plot and add tickers

plot(stocksd,stockmu,col="gray" , xlab="Standard deviation (monthly)", ylab="Average return (monthly)",xlim=c(0,1.1*max(stocksd)),ylim=c(0,1*1*max(stockmu)))
text( stocksd[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))] , 
       stockmu[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))] ,  
      labels=colnames(returns )[!is.element(names(stocksd),c("HD","CAT","XOM","AAPL"))], cex= 0.7)
text( stocksd[c("HD","CAT","XOM","AAPL")]  ,  stockmu[c("HD","CAT","XOM","AAPL")] ,  labels=  c("HD","CAT","XOM","AAPL") , cex= 0.9,col="red")
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```


--- type:NormalExercise lang:r xp:100 skills:1 key:c066b2d5f610d13afdb48831249d18142590def7
## Finding the mean-variance efficient portfolio  

A mean-variance efficient portfolio can be obtained as the solution of minimizing the portfolio variance under the constraint that the portfolio expected return equals a target return. A convenient R function for doing so is the function [portfolio.optim](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim) in the R package [tseries](http://www.rdocumentation.org/packages/tseries). 
Its default implementation finds the mean-variance efficient portfolio weights under the constraint that the portfolio return equals the return on the equally-weighted portfolio. 
The only argument needed is the monthly return data on the portfolio components for which the weights need to be determined.


*** =instructions
- load the library `tseries`. 
- define `out` as the output of applying the default version of `portfolio.optim` on `returns`.
- the optimized portfolio weights are in `out$pw`. Print them from largest to smallest, with rounding to 4 decimals.
- Visualize the weights larger than 1% through a barplot graph.
- Print the expect portfolio return (`out$pm`) and volatility (`out$ps`) of the optimized portfolio. 
- Verify that the optimized portfolio returns in `out$px` dominate the returns of the equally weighted portfolio.

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# load tseries

# optimize
out <-  
# output
optim_weights <- out$pw
names(optim_weights) <- colnames(returns)
optim_returns <-  out$px 
# round sorted
round( sort(    ,decreasing=T) ,4)
# barplot
   (optim_weights[optim_weights>=0.01])
# print expected return and volatility 

# equally weighted returns
ew_returns <- rowMeans(returns)
# strict lower standard deviation
 
# same or higher return
 
```

*** =solution
```{r}
# solution code
# load tseries. 
library(tseries)
# optimize
out <- portfolio.optim(returns)
# output
optim_weights <- out$pw
names(optim_weights) <- colnames(returns)
optim_returns <- xts(out$px,order.by=time(returns))
# round sorted
round(sort(optim_weights,decreasing=T) ,4)
# barplot
barplot(optim_weights[optim_weights>=0.01])
# dominate
ew_returns <- rowMeans(returns)
# strict lower standard deviation
out$ps ; sd(ew_returns)
( out$ps < sd(ew_returns) )
# same or higher return
out$pm ; mean(ew_returns)
out$pm >= mean(ew_returns)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1  key:2240e86100
## Effect of the return target 

A more general specification of the function [portfolio.optim](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim) is as follows:
$$
portfolio.optim(x, pm = mean(x),
                shorts = FALSE,  reshigh = NULL)
$$
The argument ``pm`` sets the target return. The argument ``shorts`` is a boolean saying whether negative weights are forbidden (``shorts=FALSE``) and the argument ``reshigh`` specifies the upper constraints on the portfolio weights. By default, when there are no negative weights allows and since the portfolio has to be fully invested, the maximum weight is 1. 

In the exercises we quantify the risk-reward trade-off when choosing the mean-variance efficient portfolio. The higher the target return, the higher will be the corresponding portfolio volatility. 

*** =instructions
- There is a isk-reward trade-off when choosing the mean-variance efficient portfolio. The higher the target return, the higher will be the corresponding portfolio volatility. When the target return is 10% higher than the average return, how much % increase is there in the portfolio volatility?  
- Compute the Sharpe ratio of the two portfolios when the monthly risk free rate is 0.1%. 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

library(tseries)
```

*** =sample_code
```{r}
meanreturns = mean(returns)
# compute standard deviation when target retun is mean return
out1 <- portfolio.optim(returns, pm = meanretuurns )
sd1 <-
 
# compute standard deviation when target retun is 1.1 times the mean return
out2 <-
sd2 <- 
# compute relative increase in standard deviation

```

*** =solution
```{r}
# solution code
# load tseries. 
library(tseries)
meanreturns = mean(returns)
# compute standard deviation when target retun is mean return
out1 <- portfolio.optim(returns, pm = meanreturns )
sd1 <- out1$psd
 
# compute standard deviation when target retun is 1.1 times the mean return
out2 <- portfolio.optim(returns, pm = 1.1*meanreturns )
sd2 <- out2$psd
# compute relative increase in standard deviation
(sd2-sd1)/sd1
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6f3df5bc7125d675f226ddc4c69639e2cf24d23c
## Imposing weight constraints

Investors are often constrained in the maximum values allowed for the portfolio weights. Finding the optimized portfolio under a maximum weight is possible in the function [portfolio.optim](http://www.rdocumentation.org/packages/tseries/functions/portfolio.optim)  by setting the argument
 `reshigh` to the vector of maximum weights for each asset.  
 
 The advantage of a maximum weight constraint is that the portfolio will be less concentrated. Let's verify this statement when the target expected return is still equal to the equally-weighted return. The disadvantage is that the same target return, may no longer be possible or will be obtained at the expense of a higher volatility. 

*** =instructions
- Modify the sample code to impose a maximum 10% constraint on each of the portfolio weights and compute the number of positions larger than 1%;
- Repeat the above with a maximum 5% constraint.
- Print the corresponding volatilities of the optimized portfolios. 

*** =hint
The argument `reshigh` needs to be a vector. 

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

library(tseries)
```

*** =sample_code
```{r}
# sample code
# tseries is pre-loaded
out <- portfolio.optim(returns,reshigh=rep(1,ncol(returns)))
sum( 1*(out$pw > 0.01))


```

*** =solution
```{r}
# solution code
out1 <- portfolio.optim(returns,reshigh=rep(1,ncol(returns)))
sum( 1*(out$pw > 0.01))

out2 <- portfolio.optim(returns,reshigh=rep(0.1,ncol(returns)))
sum( 1*(out$pw > 0.01))

out3 <- portfolio.optim(returns,reshigh=rep(0.05,ncol(returns)))
sum( 1*(out$pw > 0.01))

out1$ps
out2$ps
out3$ps
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:461436d2ee4b426f97b3c95b6ee39836ae38b34f
## The efficient frontier

Concept of minimizing variance given a sequence of return targets starting from the lowest to highest expected returns
*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v1/hls-ch4_2.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:2c03a87fc29fb66e35d8cdbf244b3dcf41305824
## Computing the efficient frontier 

As we have seen, we need to first define the grid of target returns. We will set the maximum target return to the maximum average return of the stocks. Ideally, we set the minimum target return to the return of the minimum variance portfolio. Since we don't know this minimum variance portfolio return yet, let's therefore take the largest possible grid starting with the minimum average returns of all stocks. 


*** =instructions
- Complete the code specifying the grid of 100 equally space return targets starting at the minimum and maximum average return 
- Run the loop computing the optimized weights and corresponding mean return and volatility
- Plot the portfolio mean against the portfolio volatility of the optimized portfolios
- Improve the efficient frontier plot by removing the inefficient portfolios offering a lower return at a higher risk compared to the minimum variance portfolio.  

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# sample code
# solution code
stockmu = colMeans(returns)
grid = seq( , ,length.out=100 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()
for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <- c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
colnames(mweights) = colnames(returns)
# report the scatter plot of vpm agains vpsd

```

*** =solution
```{r}
# solution code
stockmu = colMeans(returns)
grid = seq(min(stockmu),max(stockmu),length.out=100 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <- c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
# report the scatter plot of vpm agains vpsd
plot(vpms,vpm)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1   key:897a83541a
## The minimum variance and maximum Sharpe ratio portfolio

Let us investigate further the previous output consistent of the vector of expected portfolio returns (`vpm`), the 
portfolio standard deviations (`vpsd`) and the matrix of portfolio weights (`mweights`).   

 


*** =instructions
- Complete the sample code to obtain the volatility and expected return of the minimum variance portfolio. 
- Assume a risk free rate of 0.75%. Complete the sample code to compute the risk and reward of the maximum Sharpe ratio portfolio and compare its allocation with the minimum variance portfolio (focus hereby on the weights larger than 1%).
 

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
stockmu <- colMeans(returns)

grid = seq(min(stockmu),max(stockmu),length.out=100 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()

for( mutarget in grid ){
  out <- try(portfolio.optim(x = returns ,pm=mutarget),silent=TRUE)
  if(class(out)!="try-error"){
    vpm <-  c(vpm,out$pm)
    vpsd <- c(vpsd,out$ps)
    mweights <- rbind( mweights,out$pw)
  }
}
colnames(mweights) = colnames(returns)
```

*** =sample_code
```{r}
# sample code
# minimum variance portfolio
pm_minvar <- vpm[vpsd==min(vpsd)]
psd_minvar <-
weights_minvar <-
# max sharpe 
pm_maxsharpe <-  
psd_maxsharpe <-
weights_maxsharpe <-
#compare weights
par(mfrow=c(2,1), mar=c(3,2,2,1))
barplot(weights_minvar)
 
```

*** =solution
```{r}
# solution code
# minimum variance portfolio
pm_minvar <- vpm[vpsd==min(vpsd)]
psd_minvar <-vpsd[vpsd==min(vpsd)]
weights_minvar <- mweights[vpsd==min(vpsd),]
# max sharpe 
pm_maxsharpe <-  vpm[(vpm-0.0075)/vpsd==max((vpm-0.0075)/vpsd)]
psd_maxsharpe <- vpsd[(vpm-0.0075)vpsd==max((vpm-0.0075)/vpsd)]
weights_maxsharpe <- mweights[(vpm-0.0075)/vpsd==max((vpm-0.0075)/vpsd),]
par(mfrow=c(2,1), mar=c(3,2,2,1))
barplot(weights_minvar[weights_minvar>0.01])
barplot(weights_maxsharpe[weights_maxsharpe>0.01]) 
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```


--- type:MultipleChoiceExercise lang:r xp:50 skills:1    key:e647c251a7
## Properties of the efficient frontier

Consider the efficient frontier plots obtained for the DJIA stocks under a maximum 30% weight constraint (red line) and a maximum 5% weight constraint (blue). Which of the following statements is false.  

*** =instructions
- Imposing a more strict weight constraint, leads to a shift to the right of the efficient frontier.    
- The portfolios on the dashed lines are not mean-variance efficient. 
- The observed average return of any risky portfolio is always higher than the risk-free rate. 

*** =hint
hint

*** =pre_exercise_code
```{r}
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

stockmu = apply( X = returns , MARGIN = 2 , FUN="mean")
stocksd = apply( X = returns , MARGIN = 2 , FUN="sd")

grid <- seq(min(stockmu),max(stockmu),length.out=100 )  
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm = vpsd = mweights = c()

for( mutarget in grid ){
  out = try(portfolio.optim(x = returns,pm=mutarget,
                            reshigh = rep(0.3,ncol(returns)) ),silent=TRUE)
  if(class(out)!="try-error"){
    vpm = c(vpm,out$pm)
    vpsd = c(vpsd,out$ps)
    mweights = rbind( mweights,out$pw)
  }
}

# assign the stock names to the column names of mweights
colnames(mweights) = colnames(returns)

 
##################################################################################
# To avoid the concentration, we will impose now a max weight constraint of 10%

 
# Compute the efficient frontier
## for each solution, we save the portfolio mean,
## portfolio standard deviation and the weights
## Initialize them as an empty vector
vpm.res = vpsd.res = mweights.res = c()

for( mutarget in grid ){
  out = try(portfolio.optim(x = returns,pm=mutarget,
                  reshigh = rep(0.05,ncol(returns))),silent=TRUE)
  if(class(out)!="try-error"){
    vpm.res = c(vpm.res,out$pm)
    vpsd.res = c(vpsd.res,out$ps)
    mweights.res = rbind( mweights.res,out$pw)
  }
}
# assign the stock names to the column names of mweights
colnames(mweights.res) = colnames(returns)

# 
plot(stocksd,stockmu,col="gray" , xlab="Standard deviation (monthly)", ylab="Average return (monthly",
     xlim=c(0.02,0.15), ylim=c(min(stockmu),max(stockmu)))
text( stocksd ,  stockmu ,  labels=colnames(returns), cex= 0.7)
# all portfolios offering less return than the minimum variance portfolio should be excluded
minvarpm = vpm[vpsd==min(vpsd)]
minvarpm.res = vpm.res[vpsd.res==min(vpsd.res)]
lines(vpsd,vpm,col="red",lwd=1,lty=3)
lines(vpsd.res,vpm.res,col="blue",lwd=1,lty=3)
lines(vpsd[vpm>=minvarpm],vpm[vpm>=minvarpm],col="red",lwd=2)
lines(vpsd.res[vpm.res>=minvarpm.res],vpm.res[vpm.res>=minvarpm.res],col="blue",lwd=2)

```

*** =sct
```{r}
test_mc(3) # if 3 is the correct option.
```

--- type:VideoExercise lang:r xp:50 skills:1 key:ca405aabc3c21c142aaae409da6302b77b5210ef
## In-sample versus out-of-sample evaluation


*** =video_link
//player.vimeo.com/video/108225030

*** =video_hls
//videos.datacamp.com/transcoded/985_portfolio_analysis/v1/hls-ch4_3.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:69f72e3a22916ec2ef924a7f212d354f28022366
## Split-sample evaluation

The function [window](http://www.rdocumentation.org/packages/stats/functions/window), with arguments `x`, `start` 
and `end`, extracts the subperiod between `start` and `end`  from the object `x`. 

*** =instructions
- Split `returns` into the sample `returns.estim` (ending on `2003-12-31`) and `returns.eval` (starting on `2004-01-01`)
- Compute using `returns.estim` the optimal portfolio under a maximum 10% weight constraint, when the target return is the return of the equally weighted portfolio. 
- Compute using `returns.eval` the optimal portfolio under a maximum 10% weight constraint, when the target return is the return of the equally weighted portfolio.
- Make a scatter plot of the two porfolios to see how different the portfolio weights can be when changing the estimation period. Note that if portfolio weights are identical, they should be on the 45 degrees line. 
 


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]
```

*** =sample_code
```{r}
# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim = 
returns.oos = w 
# optimize portfolios
weights.estim <- portfolio.optim(x =  , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x =   ,reshigh = rep(0.1,ncol(returns.estim)))$pw
#scatterplot with 45 degrees lines
 
abline(a=0,b=1,lty=3)
```

*** =solution
```{r}
# solution code

# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim = window(returns, start = "1991-01-01", end = "2003-12-31")
returns.oos = window(returns, start ="2004-01-01", end = "2015-12-31")
# optimize portfolios
weights.estim <- portfolio.optim(x = returns.estim , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x = returns.oos ,reshigh = rep(0.1,ncol(returns.estim)))$pw
#
plot(weights.estim,weights.oos)
abline(a=0,b=1,lty=3)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6d3bdc717d87317d8c4ccf468836039b35f78440
## Out of sample performance evaluation

How severe is the optimality loss? Let's compare on the out-of-sample period the mean and variance on the optimized portfolio weights using the estimation sample and the out-of-sample evaluation data. 



*** =instructions
- Use the function [Return.portfolio](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/Return.portfolio) to compute the portfolio returns of the portfolio with weights `weights.estim` and monthly rebalancing using the `returns.estim` sample. Call the resulting portfolio returns `returns.estim` and use [table.AnnualizedReturns](http://www.rdocumentation.org/packages/PerformanceAnalytics/functions/table.AnnualizedReturns) to evaluate the expected performance based on the estimation data.
- Repeat this on the out-of-sample return data, still using `weights.estim`. This is the real performance you may expect from the optimized portfolios. 


*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(xts)
library(PerformanceAnalytics)
library(tseries)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- Return.calculate(prices, method="simple")
returns <- returns[(-1),]

# Splitting return data in sample for estimation and out of sample evaluation 
returns.estim = window(returns, start = "1991-01-01", end = "2003-12-31")
returns.oos = window(returns, start ="2004-01-01", end = "2015-12-31")
# optimize portfolios
weights.estim <- portfolio.optim(x = returns.estim , reshigh = rep(0.1,ncol(returns.estim)))$pw
weights.oos <- portfolio.optim(x = returns.oos ,reshigh = rep(0.1,ncol(returns.estim)))$pw

```

*** =sample_code
```{r}
# sample code
# weights.estim, returns.estim and returns.oos are preloaded

# portfolio returns on returns.estim
preturns.estim <- Return.portfolio( , ,rebalance_on="months")
 

# portfolio returns on returns.oos
preturns.oos <-  
 
```

*** =solution
```{r}
# solution code
# portfolio returns on returns.estim
preturns.estim <- Return.portfolio(returns.estim,weights.estim,rebalance_on="months")
table.AnnualizedReturns(preturns.estim)
# portfolio returns on returns.oos
preturns.oos <- Return.portfolio(returns.oos,weights.estim,rebalance_on="months")
table.AnnualizedReturns(preturns.oos)
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

 
--- type:NormalExercise lang:r xp:100 skills:1   key:9d55676210
## It ain't over
 
You made it! This journey showed you how to compute returns, analyze portfolio performance and optimize portfolios. These are important skills, since whenever you hold or manage more than one asset, you have a portfolio view to consider. 

It is now time to compound and get extra return by exploring other useful R packages for portfolio analysis like [PortfolioAnalytics](). Instead of optimizing the portfolio by minimizing the variance under a return targetg constraint, this package allows you optimize virtually all objective functions under all possible constraints. Even if the solution to the problem is not feasible, it will give you the best possible solution using heuristics such as differential evoluation in the R package [DEoptim](). Since I have been involved in both packages, this last exercise is an act of shameless selfpromotion. 

Have fun! 



*** =instructions
- Load the package PortfolioAnalytics and explore the functionality of this package by entering the instruction `?PortfolioAnalytics`. 
- Read my article on portfolio optimization with DEoptim. 
*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
library(Biobase)
uri <- "https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Ardia~et~al.pdf"
download.file(uri,"deoptim.pdf",method="internal",mode="wb")

```

*** =sample_code
```{r}
# load and explore the package PortfolioAnalytics

# happy reading!
openPDF("deoptim.pdf")
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```



