---
title       : Optimizing the portfolio 
description : We have up to now considered the portfolio weights as given. Here we consider optimizing the portfolio weights. Markowitz mean variance optimized portfolios.


--- type:MultipleChoiceExercise lang:r xp:50 skills:1   key:139d7eabbd827860709b29f2a8d968614d1e3cb0
## Mean variance efficient portfolio

In the workspace, you see a scatter plot showing the mean and standard deviation of the 30 Dow Jones Industrial Average stocks, as estimated on their monthly returns over the period 1991-2015. Each of the stocks is identified in the plot using
its ticker. 

Assume that 

*** =instructions
- option 1
- option 2
- option 3

*** =hint
hint

*** =pre_exercise_code
```{r}
# pec
# pec
library(xts)
library(PerformanceAnalytics)
download.file("http://s3.amazonaws.com/assets.datacamp.com/course/portfolio-analysis/prices.rds", destfile="prices.rds")
# load("prices.Rdata")
prices<- readRDS("prices.rds") 
returns <- CalculateReturns(prices, method="simple")
returns <- returns[(-1),]
```

*** =sct
```{r}
test_mc(2) # if 2 is the correct option.
```



--- type:VideoExercise lang:r xp:50 skills:1 key:2d45c303bc2e49767a9cd744f15a31938a2ec7d9
## Modern portfolio theory of Harry Markowitz

Concept of minimizing variance given a return target
*** =video_link
//player.vimeo.com/video/108225030



--- type:NormalExercise lang:r xp:100 skills:1 key:c066b2d5f610d13afdb48831249d18142590def7
## Find the mean-variance efficient portfolio with target equal to EW (step i)

Obtain the optimized portfolio weights using portfolio.optim(), plot the portfolio weights using barplot()

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6f3df5bc7125d675f226ddc4c69639e2cf24d23c
## Find the mean-variance efficient portfolio with target equal to EW (step ii)

Interpretation of the output (achieved mean return, portfolio variance).

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:461436d2ee4b426f97b3c95b6ee39836ae38b34f
## Efficient frontier

Concept of minimizing variance given a sequence of return targets starting from the lowest to highest expected returns
*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:2c03a87fc29fb66e35d8cdbf244b3dcf41305824
## Plotting the efficient frontier in R using a loop (step i)

Define the grid of target returns. Initialize the portfolio weights, means and sd to be saved.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:5cd26622b81e29df3e9a1636d88d48ed0ae3c3c5
## Plotting the efficient frontier in R using a loop (step ii)

Write the loop. for(), portfolio.optim()

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:c0483cbbf9a142a2e4541614bbeb9b8b407328d5
## plotting the efficient frontier in R using a loop (step iii)

WPlot the results

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:ca405aabc3c21c142aaae409da6302b77b5210ef
## In-sample versus out-of-sample evalution


*** =video_link
//player.vimeo.com/video/108225030

--- type:NormalExercise lang:r xp:100 skills:1 key:69f72e3a22916ec2ef924a7f212d354f28022366
## Split sample in two and compare realized returns and volatilities of assets with predicted ones

Estimation window, out of sample evaluation window. Estimation error

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:6d3bdc717d87317d8c4ccf468836039b35f78440
## Split sample in two and compare EW vs Inverse vol vs mean variance optimized portfolio

Estimation window, out of sample evaluation window. Compare in sample estimated performance with out of sample realized performance.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# pec
```

*** =sample_code
```{r}
# sample code
```

*** =solution
```{r}
# solution code
```

*** =sct
```{r}
# sct code
success_msg("Well done!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:75eb737c24e2ba235ee1de43c66a6324a9282b08
## Main take-aways

summary and extensions.
*** =video_link
//player.vimeo.com/video/108225030

